---
title: "Fragmentomics analysis"
author: "Miriam Fuentes"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    code_folding: show
    theme: cosmo
editor_options: 
  markdown: 
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
#Chunk options
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,fig.width = 8,fig.height = 5,fig.align = "center")

#Packages
library(knitr)
library(tidyverse)
library(data.table)
library(patchwork)
timepoint_colours<-c("BL"="darkcyan","END"="lightpink4")

timepoint_labels<-c("BL"="Baseline","END"="After radiotherapy")

common_theme<-theme_classic(base_size=15)+
  theme(
    plot.title=element_text(size=17,face="bold"),
    axis.title.y=element_text(size=16,face="bold"),
    axis.text.y=element_text(size=13,face="bold"),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    legend.title=element_text(size=14,face="bold"),
    legend.text=element_text(size=13),
    legend.position="right",
    aspect.ratio=1)
```

# 1 Fragment-count preprocessing

## 1.1 Extract fragment sizes

```bash
analysis_root="../../nextflow_integrator/devonvolutor_dedup/deconvolutor_dedup/deconvolutor_analysis"
mkdir -p "$analysis_root/scripts"
cd "$analysis_root"

cat > scripts/get_sizes.py <<'PY'
import pysam, glob, os
from collections import Counter

path="../../../results_mapped_dedup/dedup_mapped"
out="counts"

os.makedirs(out,exist_ok=True)

for bam in glob.glob(path+"/*_dedup.bam"):
    counts=Counter()

    with pysam.AlignmentFile(bam,"rb") as f:
        for read in f:
            if not read.is_read1: continue
            if not read.is_paired: continue
            if read.is_unmapped or read.mate_is_unmapped: continue
            if read.is_secondary or read.is_supplementary: continue
            if read.is_duplicate: continue
            if read.mapping_quality<=30: continue

            size=abs(read.template_length)

            if 50<=size<=250:
                counts[size]+=1

    name=os.path.basename(bam).replace("_dedup.bam","")

    with open(out+"/"+name+"_size.txt","w") as f:
        for size in range(50,251):
            f.write(f"{size}\t{counts[size]}\n")

    print(name,sum(counts.values()))
PY
```

## 1.2 Run deconvolution

```bash
cat > run_deconv.sh <<'BASH'
#!/bin/bash
analysis_root="../../nextflow_integrator/devonvolutor_dedup/deconvolutor_dedup/deconvolutor_analysis"
cd "$analysis_root"

counts="counts"
results="results"

mkdir -p "$results"

echo "sample,amplitude_ratio,entropy_ratio" > "$results/deconvolution_results.csv"

for file in "$counts"/*_size.txt; do

    name=$(basename "$file" _size.txt)

    cp "$file" example_size.txt
    cp "$file" example_size_sWGS.txt

    python size_decov.py > output.txt

    amp=$(grep "amplitude ratio" output.txt | tail -1 | awk '{print $NF}')
    ent=$(grep "entropy ratio" output.txt | tail -1 | awk '{print $NF}')

    echo "$name,$amp,$ent" >> "$results/deconvolution_results.csv"

    echo "$name done"

done
BASH
```

# 2 Load and prepare data

## 2.1 Paths and settings

```{r paths-and-settings}
project_path <- file.path("..","..","nextflow_integrator")
root <- file.path(project_path,"devonvolutor_dedup","deconvolutor_dedup","deconvolutor_analysis")

results_path <- file.path(root, "results")
count_path <- file.path(root, "counts")

peaks <- c(121,131,141,151,159,167,177,188,199)
```

## 2.2 Load ratios, counts and components

```{r load-fragmentomics-data}
#Load ratios and fragment counts
results <- fread(file.path(results_path,"deconvolution_ratios.csv")) %>%
  filter(model=="constrained_L2") %>%
  mutate(
    patient=sub("_(BL|END)$","",sample),
    timepoint=sub("^.*_","",sample),
    timepoint=factor(timepoint,levels=c("BL","END")))

files <- list.files(count_path,pattern="_size.txt$",full.names=TRUE)

size_data <- lapply(files,function(f) {
  x <- fread(f,col.names=c("size","count"))
  x$sample <- sub("_size.txt","",basename(f))
  x
}) %>%
  bind_rows() %>%
  mutate(patient=sub("_(BL|END)$","",sample),timepoint=sub("^.*_","",sample))

#Load fitted components
components <- fread(file.path(results_path,"deconvolution_components_L2.csv")) %>%
  arrange(sample,location) %>%
  group_by(sample) %>%
  mutate(component=peaks[row_number()]) %>%
  ungroup() %>%
  mutate(patient=sub("_(BL|END)$","",sample),timepoint=sub("^.*_","",sample))
table(results$timepoint)
length(unique(size_data$sample))
table(components$component)
```

# 3 Main analysis

## 3.1 Pooled fragment-size profile

```{r pooled-fragment-profile}
profile <- size_data %>%
  group_by(timepoint,size) %>%
  summarise(count=sum(count),.groups="drop") %>%
  group_by(timepoint) %>%
  mutate(freq=100*count/sum(count))
ggplot(profile,aes(size,freq,colour=timepoint)) +
  geom_line(linewidth=1) +
  scale_colour_manual(values=timepoint_colours,labels=timepoint_labels) +
  labs(x="Fragment size (bp)",y="Frequency (%)",colour="") +
  theme_classic(base_size=15)
```

## 3.2 Global deconvolution ratios

```{r global-deconvolution-ratios}
paired <- results %>%
  select(patient,timepoint,amplitude_ratio,entropy_ratio) %>%
  pivot_wider(names_from=timepoint,values_from=c(amplitude_ratio,entropy_ratio))

results %>%
  group_by(timepoint) %>%
  summarise(n=n(),amplitude=median(amplitude_ratio),entropy=median(entropy_ratio))
amp_ratio_test <- wilcox.test(paired$amplitude_ratio_END,paired$amplitude_ratio_BL,paired=TRUE,exact=FALSE)

entropy_ratio_test <- wilcox.test(paired$entropy_ratio_END,paired$entropy_ratio_BL,paired=TRUE,exact=FALSE)

amp_ratio_test
entropy_ratio_test
```

# 4 Secondary analyses

## 4.1 Fragment-size metrics

```{r fragment-size-metrics}
frag <- size_data %>%
  group_by(patient,timepoint) %>%
  summarise(
    short150=100*sum(count[size<150])/sum(count),
    short159=100*sum(count[size<159])/sum(count),
    short_long=sum(count[size>=100 & size<=150])/
               sum(count[size>=151 & size<=220]),
    .groups="drop") %>%
  pivot_wider(names_from=timepoint,values_from=c(short150,short159,short_long))

frag_tests <- data.frame(
  metric=c("<150 bp","<159 bp","Short/long ratio"),
  BL=c(median(frag$short150_BL),median(frag$short159_BL),median(frag$short_long_BL)),
  END=c(median(frag$short150_END),median(frag$short159_END),median(frag$short_long_END)),
  change=c(
    median(frag$short150_END-frag$short150_BL),
    median(frag$short159_END-frag$short159_BL),
    median(frag$short_long_END-frag$short_long_BL)),
  p=c(
    wilcox.test(frag$short150_END,frag$short150_BL,paired=TRUE,exact=FALSE)$p.value,
    wilcox.test(frag$short159_END,frag$short159_BL,paired=TRUE,exact=FALSE)$p.value,
    wilcox.test(frag$short_long_END,frag$short_long_BL,
                paired=TRUE,exact=FALSE)$p.value))

frag_tests$p_adj <- p.adjust(frag_tests$p,"BH")

frag_tests
```

## 4.2 Component amplitudes

```{r component-amplitude-tests}
amp_tests <- data.frame()

for (p in peaks) {
  x <- components %>%
    filter(component==p) %>%
    select(patient,timepoint,amplitude) %>%
    pivot_wider(names_from=timepoint,values_from=amplitude)

  test <- wilcox.test(x$END,x$BL,paired=TRUE,exact=FALSE)

  amp_tests <- rbind(amp_tests,data.frame(
    component=p,
    BL=median(x$BL),
    END=median(x$END),
    change=median(x$END-x$BL),
    p=test$p.value))
}

amp_tests$p_adj <- p.adjust(amp_tests$p,"BH")

amp_tests
amp_tests %>% filter(p_adj<0.05)
```

## 4.3 Component entropy

### 4.3.1 Prepare entropy data

```{r prepare-component-entropy}
components <- components %>%
  mutate(
    entropy=log(4*pi*scale),
    entropy_group=case_when(
      component %in% c(121,131) ~ "121/131",
      component %in% c(141,151) ~ "141/151",
      component==159 ~ "159",
      component>=167 ~ "167+"))
entropy_data <- components %>%
  distinct(sample,patient,timepoint,entropy_group,.keep_all=TRUE)
```

### 4.3.2 Test entropy changes

```{r component-entropy-tests}
groups <- c("121/131","141/151","159","167+")

entropy_tests <- data.frame()

for (g in groups) {
  x <- entropy_data %>%
    filter(entropy_group==g) %>%
    select(patient,timepoint,entropy) %>%
    pivot_wider(names_from=timepoint,values_from=entropy)

  test <- wilcox.test(x$END,x$BL,paired=TRUE,exact=FALSE)

  entropy_tests <- rbind(entropy_tests,data.frame(
    group=g,
    BL=median(x$BL),
    END=median(x$END),
    change=median(x$END-x$BL),
    p=test$p.value))
}

entropy_tests$p_adj <- p.adjust(entropy_tests$p,"BH")

entropy_tests
```

## 4.4 Fragment-size-specific changes

```{r fragment-size-tests}
size_freq <- size_data %>%
  group_by(sample,patient,timepoint) %>%
  mutate(freq=100*count/sum(count)) %>%
  ungroup()

size_paired <- size_freq %>%
  select(patient,timepoint,size,freq) %>%
  pivot_wider(names_from=timepoint,values_from=freq)
size_tests <- size_paired %>%
  group_by(size) %>%
  summarise(
    BL_median=median(BL),
    END_median=median(END),
    change=median(END-BL),
    p=wilcox.test(END,BL,paired=TRUE,exact=FALSE)$p.value,
    .groups="drop") %>%
  mutate(p_adj=p.adjust(p,"BH"))
size_tests %>%
  arrange(p_adj,p) %>%
  head(20)
size_tests %>%
  filter(size>=115 & size<=135)
```

## 4.5 Fragments 115-125 bp

```{r fragment-window-analysis}
region121 <- size_data %>%
  group_by(patient,timepoint) %>%
  summarise(percent_115_125= 100*sum(count[size>=115 & size<=125])/sum(count),.groups="drop") %>%
  pivot_wider(names_from=timepoint,values_from=percent_115_125)
wilcox.test(region121$END,region121$BL,paired=TRUE,exact=FALSE)
region121 %>%
  summarise(BL=median(BL),END=median(END),change=median(END-BL))
```

# 5 Figures

## 5.1 Build figure panels

```{r build-fragmentomics-figures}
figure_path <- file.path(project_path,"fragmentomics","figures")
dir.create(figure_path,recursive=TRUE,showWarnings=FALSE)

plot_colours <- c("BL"="darkcyan","END"="lightpink4")
plot_labels <- c("BL"="Baseline","END"="Post-RT")

#Panel A: component amplitudes
main_A <- ggplot(amp_tests,aes(factor(component),change))+
  geom_hline(yintercept=0,linetype="dashed",colour="grey60")+
  geom_segment(aes(xend=factor(component),y=0,yend=change),colour="grey70")+
  geom_point(shape=21,fill="white",colour="grey50",size=2)+
  geom_point(data=amp_tests %>% filter(component==121),shape=21,fill="darkcyan",colour="black",size=2.5)+
  labs(x="Fitted component (bp)",y="Amplitude change\n(Post-RT - baseline)")+
  theme_classic(base_size=10)

#Panel B: amplitude ratio
main_B <- ggplot(results,aes(timepoint,amplitude_ratio,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_line(aes(group=patient),colour="grey80",linewidth=0.3)+
  geom_point(size=1.3)+
  scale_fill_manual(values=plot_colours)+
  scale_x_discrete(labels=plot_labels)+
  labs(x=NULL,y="Amplitude ratio",title=paste0("Wilcoxon p = ",sprintf("%.3f",amp_ratio_test$p.value)))+
  theme_classic(base_size=10)+
  theme(legend.position="none",plot.title=element_text(size=8,hjust=0.5))

#Panel C: entropy ratio
main_C <- ggplot(results,aes(timepoint,entropy_ratio,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_line(aes(group=patient),colour="grey80",linewidth=0.3)+
  geom_point(size=1.3)+
  scale_fill_manual(values=plot_colours)+
  scale_x_discrete(labels=plot_labels)+
  labs(x=NULL,y="Entropy ratio",title=paste0("Wilcoxon p = ",sprintf("%.3f",entropy_ratio_test$p.value)))+
  theme_classic(base_size=10)+
  theme(legend.position="none",plot.title=element_text(size=8,hjust=0.5))

#Panel D: 121-bp amplitude
component_121 <- components %>%
  filter(component==121) %>%
  mutate(timepoint=factor(timepoint,levels=c("BL","END")))

p121 <- amp_tests$p_adj[amp_tests$component==121]

main_D <- ggplot(component_121,aes(timepoint,amplitude,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_line(aes(group=patient),colour="grey80",linewidth=0.3)+
  geom_point(size=1.3)+
  scale_fill_manual(values=plot_colours)+
  scale_x_discrete(labels=plot_labels)+
  labs(x=NULL,y="121-bp amplitude (%)",title=paste0("BH-adjusted p = ",sprintf("%.3f",p121)))+
  theme_classic(base_size=10)+
  theme(legend.position="none",plot.title=element_text(size=8,hjust=0.5))

#Panel E: fragments 115-125 bp
region121_plot <- region121 %>%
  pivot_longer(c(BL,END),names_to="timepoint",values_to="percent") %>%
  mutate(timepoint=factor(timepoint,levels=c("BL","END")))

region121_test <- wilcox.test(region121$END,region121$BL,paired=TRUE,exact=FALSE)

main_E <- ggplot(region121_plot,aes(timepoint,percent,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_line(aes(group=patient),colour="grey80",linewidth=0.3)+
  geom_point(size=1.3)+
  scale_fill_manual(values=plot_colours)+
  scale_x_discrete(labels=plot_labels)+
  labs(x=NULL,y="Fragments 115-125 bp (%)",title=paste0("Wilcoxon p = ",sprintf("%.3f",region121_test$p.value)))+
  theme_classic(base_size=10)+
  theme(legend.position="none",plot.title=element_text(size=8,hjust=0.5))

#Combine panels
fragmentomics_figure <- main_A /
  (main_B | main_C) /
  (main_D | main_E) +
  plot_layout(heights=c(0.8,1,1))+
  plot_annotation(tag_levels="A")

fragmentomics_figure

#Save figure
ggsave(file.path(figure_path,"Figure_fragmentomics.png"),fragmentomics_figure,width=10,height=9,dpi=600,bg="white")

ggsave(file.path(figure_path,"Figure_fragmentomics.pdf"),fragmentomics_figure,width=10,height=9,bg="white")
```

## 5.2 Main figure

```{r main-figure, fig.width=11, fig.height=9.5}
main_panel <- (main_B | main_C) /
  main_A /
  (main_D | main_E) +
  plot_layout(heights=c(1.15,0.70,1.15),guides="collect") +
  plot_annotation(
    tag_levels="A",
    theme=theme(
      plot.tag=element_text(size=13,face="bold"),
      plot.background=element_rect(colour="black",fill="white",linewidth=0.8),
      plot.margin=margin(8,8,8,8))) &
  theme(
    axis.title=element_text(size=10,face="bold"),
    axis.text=element_text(size=8.5,colour="black"),
    plot.title=element_text(size=9.5,face="plain",hjust=0.5,margin=margin(b=5)),
    plot.margin=margin(6,9,6,9),
    legend.position="bottom",
    legend.direction="horizontal",
    legend.title=element_blank(),
    legend.text=element_text(size=9),
    legend.key.width=grid::unit(8,"mm"),
    legend.key.height=grid::unit(3.5,"mm"))

main_panel

ggsave(file.path(results_path,"main_fragmentomics.pdf"),main_panel,width=11,height=9.5,units="in",bg="white")

ggsave(file.path(results_path,"main_fragmentomics.png"),main_panel,width=11,height=9.5,units="in",dpi=400,bg="white")
```

## 5.3 Supplementary figure

```{r supplementary-figure, fig.width=11, fig.height=7.5}
supp_theme <- theme_classic(base_size=10.5) +
  theme(
    axis.title=element_text(size=10,face="bold"),
    axis.text=element_text(size=8.5,colour="black"),
    plot.title=element_text(size=9.5,face="plain",hjust=0.5),
    plot.margin=margin(7,9,7,9))

#Panel A: fragment-size profile
supp_A <- ggplot(profile,aes(size,freq,colour=timepoint)) +
  geom_line(linewidth=0.8) +
  scale_colour_manual(
    values=timepoint_colours,
    breaks=c("BL","END"),
    labels=c("BL"="Baseline","END"="Post-RT"),
    name=NULL) +
  guides(colour=guide_legend(nrow=1)) +
  labs(x="Fragment size (bp)",y="Frequency (%)") +
  supp_theme

#Panel B: entropy changes
entropy_plot_df <- entropy_tests %>%
  mutate(group=factor(group,levels=c("121/131","141/151","159","167+")))

supp_B <- ggplot(entropy_plot_df,aes(group,change)) +
  geom_hline(yintercept=0,linetype="dashed",colour="grey45",linewidth=0.4) +
  geom_segment(aes(xend=group,y=0,yend=change),colour="grey70",linewidth=0.55) +
  geom_point(shape=21,fill="grey75",colour="black",size=2.3,stroke=0.35) +
  scale_y_continuous(n.breaks=5,expand=expansion(mult=c(0.15,0.15))) +
  labs(x="Component group (bp)",y="Median entropy change\n(Post-RT - baseline)") +
  supp_theme

#Panel C: fragment-size changes
size_change <- size_tests %>%
  filter(size>=100,size<=140)

supp_C <- ggplot(size_change,aes(size,change)) +
  annotate("rect",xmin=115,xmax=125,ymin=-Inf,ymax=Inf,fill="grey88") +
  geom_hline(yintercept=0,linetype="dashed",colour="grey45",linewidth=0.4) +
  geom_segment(aes(xend=size,y=0,yend=change),colour="grey70",linewidth=0.35) +
  geom_point(data=size_change %>% filter(p_adj>=0.05),shape=21,fill="grey75",colour="grey45",size=1.7,stroke=0.25) +
  geom_point(data=size_change %>% filter(p_adj<0.05),shape=21,fill="black",colour="black",size=2.1,stroke=0.25) +
  scale_x_continuous(breaks=seq(100,140,5)) +
  scale_y_continuous(n.breaks=5,expand=expansion(mult=c(0.12,0.12))) +
  labs(x="Fragment size (bp)",y="Median frequency change\n(Post-RT - baseline, %)") +
  supp_theme

#Combine panels
supp_panel <- supp_A /
  (supp_B | supp_C) +
  plot_layout(heights=c(1.05,1),guides="collect") +
  plot_annotation(
    tag_levels="A",
    theme=theme(
      plot.tag=element_text(size=13,face="bold"),
      plot.background=element_rect(colour="black",fill="white",linewidth=0.8),
      plot.margin=margin(8,8,8,8))) &
  theme(
    legend.position="bottom",
    legend.direction="horizontal",
    legend.title=element_blank(),
    legend.text=element_text(size=9),
    legend.key.width=grid::unit(8,"mm"),
    legend.key.height=grid::unit(3.5,"mm"))

supp_panel

ggsave(file.path(results_path,"supplementary_fragmentomics.pdf"),supp_panel,width=11,height=7.5,units="in",bg="white")

ggsave(file.path(results_path,"supplementary_fragmentomics.png"),supp_panel,width=11,height=7.5,units="in",dpi=400,bg="white")
```

## 5.4 Combined fragmentomics panel

```{r full-fragmentomics-panel-compact, fig.width=10, fig.height=9}
design <- "
AAABBB
CCDDEE
#FFGG#
"

full_panel <- wrap_plots(supp_A,main_A,main_B,main_C,supp_B,main_D,main_E,design = design,guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    axis.title = element_text(size = 9, face = "bold"),
    axis.text = element_text(size = 7.5, colour = "black"),
    plot.title = element_text(size = 8, hjust = 0.5),
    plot.margin = margin(4, 5, 4, 5),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 8))

full_panel
```

## 5.5 Save combined panel

```{r save-full-fragmentomics-panel}
ggsave(file.path(figure_path,"fragmentomics_full_panel.pdf"),full_panel,width=11,height=8,units="in",bg="white")

ggsave(file.path(figure_path,"fragmentomics_full_panel.png"),full_panel,width=11,height=8,units="in",dpi=400,bg="white")
```
