---
title: "Targeted-panel sequencing QC"
author: "Miriam Fuentes"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    code_folding: show
    theme: cosmo
editor_options: 
  markdown: 
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
#Chunk options
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,fig.width = 8,fig.height = 5,fig.align = "center")

#Packages
library(knitr)
library(tidyverse)
library(DT)
library(writexl)
library(ggVennDiagram)
library(pheatmap)
library(wesanderson)
library(ggnewscale)
library(scales)
library(gt)
library(patchwork)
library(viridis)

#Figure style
source(file.path("..","..","nextflow_integrator","unified_figures.R"))
```

# 1 Sequencing yield

## 1.1 Prepare sequencing-yield input

```bash
mkdir -p qc_thesis

awk -F '\t' '
BEGIN {OFS="\t"}
NR==1 || $2!="1452"
' qc_thesis/sequencing_yield.tsv \
> qc_thesis/sequencing_yield_thesis.tsv

echo "Samples:"
awk 'END {print NR-1}' qc_thesis/sequencing_yield_thesis.tsv

echo "Patient 1452 rows:"
awk -F '\t' '$2=="1452"' qc_thesis/sequencing_yield_thesis.tsv
```

## 1.2 Load sequencing-yield data

```{r load-sequencing-yield}
path <- file.path("..","..","nextflow_integrator","multiqc","qc_data")

list.files(path)

qc_yield <- read.delim(file.path(path,"sequencing_yield_thesis.tsv"),sep="\t",header=TRUE,colClasses="character")

dim(qc_yield)
head(qc_yield)
```

## 1.3 Prepare sequencing-yield data

```{r prepare-sequencing-yield}
yield_data <- qc_yield %>%
  filter(patient!="1452") %>%
  mutate(read_pairs=as.numeric(read_pairs),read_pairs_million=read_pairs/1000000,sample_group="Patient samples")
yield_data %>%
  count(sample_group)
yield_summary <- yield_data %>%
  summarise(
    samples=n(),
    median=median(read_pairs_million,na.rm=TRUE),
    IQR=IQR(read_pairs_million,na.rm=TRUE),
    minimum=min(read_pairs_million,na.rm=TRUE),
    maximum=max(read_pairs_million,na.rm=TRUE))

yield_summary
```

## 1.4 Sequencing-yield plot

```{r plot-sequencing-yield}
figure_1<-ggplot(yield_data,aes(x=sample_group,y=read_pairs_million))+
  geom_boxplot(width=0.4,outlier.shape=NA,fill="darkcyan",alpha=0.7)+
  geom_jitter(width=0.08,size=2,alpha=0.7)+
  labs(x="",y="Sequencing yield (million read pairs)")+
  theme_classic(base_size=14)

figure_1
```

## 1.5 Mapping performance

```{r mapping-performance}
mapping_summary<-qc_yield %>%
    filter(patient!="1452") %>%
    mutate(mapped_percent=as.numeric(mapped_percent),properly_paired_percent=as.numeric(properly_paired_percent)) %>%
    summarise(
      samples=n(),
      mapped_median=median(mapped_percent,na.rm=TRUE),
      mapped_Q1=quantile(mapped_percent,0.25,na.rm=TRUE),
      mapped_Q3=quantile(mapped_percent,0.75,na.rm=TRUE),
      mapped_minimum=min(mapped_percent,na.rm=TRUE),
      mapped_maximum=max(mapped_percent,na.rm=TRUE),
      properly_paired_median=median(properly_paired_percent,na.rm=TRUE),
      properly_paired_Q1=quantile(properly_paired_percent,0.25,na.rm=TRUE),
      properly_paired_Q3=quantile(properly_paired_percent,0.75,na.rm=TRUE),
      properly_paired_minimum=min(properly_paired_percent,na.rm=TRUE),
      properly_paired_maximum=max(properly_paired_percent,na.rm=TRUE))

mapping_table <- tibble(
  Metric=c("Mapped reads","Properly paired reads"),
  Median=c(mapping_summary$mapped_median,mapping_summary$properly_paired_median),
  `IQR (Q1-Q3)`=c(
    sprintf("%.2f-%.2f",mapping_summary$mapped_Q1,mapping_summary$mapped_Q3),
    sprintf(
      "%.2f-%.2f",
      mapping_summary$properly_paired_Q1,
      mapping_summary$properly_paired_Q3)),
  Range=c(
    sprintf("%.2f-%.2f",mapping_summary$mapped_minimum,mapping_summary$mapped_maximum),
    sprintf(
      "%.2f-%.2f",
      mapping_summary$properly_paired_minimum,
      mapping_summary$properly_paired_maximum)))

mapping_gt<-mapping_table %>%
    gt() %>%
    tab_header(title=md("**Raw-read mapping performance**"),subtitle="Patient plasma samples (n = 104)") %>%
    cols_label(Metric="QC metric",Median="Median (%)",`IQR (Q1-Q3)`="IQR (%)",Range="Range (%)") %>%
    fmt_number(columns=Median,decimals=2) %>%
    cols_align(align="left",columns=Metric) %>%
    cols_align(align="center",columns=c(Median,`IQR (Q1-Q3)`,Range)) %>%
    tab_style(style=cell_text(weight="bold"),locations=cells_column_labels()) %>%
    tab_style(style=cell_text(weight="bold"),locations=cells_body(columns=Metric)) %>%
    tab_source_note(source_note=md("*Values were calculated from the raw mapped BAM files.*")) %>%
    tab_options(
      table.font.size=11,
      heading.align="left",
      heading.title.font.size=14,
      heading.subtitle.font.size=11,
      column_labels.background.color="#DCE6EF",
      column_labels.padding=6,
      data_row.padding=6,
      table.border.top.color="#334E68",
      table.border.top.width=px(2),
      table.border.bottom.color="#334E68",
      table.border.bottom.width=px(2))

  mapping_gt
```

# 2 Mean target coverage

## 2.1 Prepare HsMetrics input

```bash
ROOT=results_bam_to_dx_batches
ALL=qc_thesis/hsmetrics.tsv
THESIS=qc_thesis/hsmetrics_thesis.tsv

printf 'sample\tpatient\tendpoint\tstage\tmean_target_coverage\tmedian_target_coverage\tpct_selected_bases\tfold_80_base_penalty\tzero_cvg_targets_pct\tpct_target_bases_10x\tpct_target_bases_20x\tpct_target_bases_50x\tpct_target_bases_100x\tpct_target_bases_250x\tpct_target_bases_500x\tpct_target_bases_1000x\n' > "$ALL"

for file in $(find "$ROOT" -type f \( \
    -name '*_raw_mapped.CollectHsMetrics.coverage_metrics' -o \
    -name '*_dx_mapped.CollectHsMetrics.coverage_metrics' \)); do

    name=$(basename "$file")

    if [[ "$name" == *_raw_mapped.CollectHsMetrics.coverage_metrics ]]; then
        sample=${name%_raw_mapped.CollectHsMetrics.coverage_metrics}
        stage="Raw"
    else
        sample=${name%_dx_mapped.CollectHsMetrics.coverage_metrics}
        stage="Duplex"
    fi

    patient=${sample%_*}
    endpoint=${sample##*_}

    awk -F '\t' \
        -v sample="$sample" \
        -v patient="$patient" \
        -v endpoint="$endpoint" \
        -v stage="$stage" '
    BEGIN {OFS="\t"}

    /^#/ || NF==0 {next}

    header==0 {
        for(i=1;i<=NF;i++) column[$i]=i
        header=1
        next
    }

    header==1 {
        print sample,
              patient,
              endpoint,
              stage,
              get_value("MEAN_TARGET_COVERAGE"),
              get_value("MEDIAN_TARGET_COVERAGE"),
              get_value("PCT_SELECTED_BASES"),
              get_value("FOLD_80_BASE_PENALTY"),
              get_value("ZERO_CVG_TARGETS_PCT"),
              get_value("PCT_TARGET_BASES_10X"),
              get_value("PCT_TARGET_BASES_20X"),
              get_value("PCT_TARGET_BASES_50X"),
              get_value("PCT_TARGET_BASES_100X"),
              get_value("PCT_TARGET_BASES_250X"),
              get_value("PCT_TARGET_BASES_500X"),
              get_value("PCT_TARGET_BASES_1000X")
        exit
    }

    function get_value(name) {
        if(name in column) return $(column[name])
        return "NA"
    }' "$file" >> "$ALL"
done

#Remove patient 1452
awk -F '\t' '
BEGIN {OFS="\t"}
NR==1 || $2!="1452"
' "$ALL" > "$THESIS"

echo "HsMetrics counts:"

awk -F '\t' '
NR>1 {count[$4]++}
END {
    print "Raw:",count["Raw"]
    print "Duplex:",count["Duplex"]
    print "Total:",NR-1
}' "$THESIS"

echo
echo "First rows:"
head -n 5 "$THESIS"

echo
ls -lh "$THESIS"
```

## 2.2 Load HsMetrics

```{r load-hsmetrics}
hsmetrics_file <- file.path(path,"hsmetrics.tsv")

hsmetrics <- read.delim(hsmetrics_file,sep="\t",header=TRUE,colClasses="character")

hsmetrics <- hsmetrics %>%
  filter(patient!="1452")

dim(hsmetrics)

hsmetrics %>%
  count(stage)
```

## 2.3 Coverage summary

```{r prepare-mean-coverage}
coverage_data <- hsmetrics %>%
  mutate(
    mean_target_coverage=as.numeric(mean_target_coverage),
    stage=factor(
      stage,
      levels=c("Raw","Duplex"),
      labels=c("Raw reads","Duplex consensus")))
sum(is.na(coverage_data$mean_target_coverage))
coverage_data %>%
  count(sample) %>%
  filter(n!=2)
```

```{r summarise-mean-coverage}
coverage_summary <- coverage_data %>%
  group_by(stage) %>%
  summarise(
    samples=n(),
    median=median(mean_target_coverage,na.rm=TRUE),
    IQR=IQR(mean_target_coverage,na.rm=TRUE),
    minimum=min(mean_target_coverage,na.rm=TRUE),
    maximum=max(mean_target_coverage,na.rm=TRUE),
    .groups="drop")

coverage_summary

coverage_summary %>%
  mutate(across(c(median,IQR,minimum,maximum),~round(.x,1)))
```

## 2.4 Coverage plot

```{r plot-mean-coverage}
coverage_colours<-c("Raw reads"="grey65","Duplex consensus"="darkcyan")

figure_2<-ggplot(coverage_data,aes(x=stage,y=mean_target_coverage,fill=stage))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_jitter(width=0.08,size=2,alpha=0.7)+
  scale_fill_manual(values=coverage_colours)+
  scale_y_log10(labels=scales::label_comma())+
  labs(x="",y="Mean target coverage (×, log scale)")+
  theme_classic(base_size=14)+
  theme(legend.position="none")

figure_2
```

# 3 Target coverage breadth

## 3.1 Prepare data

```{r inspect-breadth-columns}
names(hsmetrics)[grepl("^pct_target_bases_",names(hsmetrics))]

breadth_data <- hsmetrics %>%
  mutate(
    stage=factor(stage,levels=c("Raw","Duplex"),labels=c("Raw reads","Duplex consensus")),
    across(
      starts_with("pct_target_bases_"),
      as.numeric))
breadth_data %>%
  count(stage)
```

```{r prepare-coverage-breadth}
breadth_long <- hsmetrics %>%
  mutate(
    stage=factor(stage,levels=c("Raw","Duplex"),labels=c("Raw reads","Duplex consensus")),
    across(
      starts_with("pct_target_bases_"),
      as.numeric)) %>%
  select(sample,stage,starts_with("pct_target_bases_")) %>%
  pivot_longer(cols=starts_with("pct_target_bases_"),names_to="threshold",values_to="breadth") %>%
  mutate(coverage_threshold=parse_number(threshold),breadth_percent=breadth*100)
dim(breadth_long)
breadth_long %>%count(stage,coverage_threshold)
head(breadth_long)
range(breadth_long$breadth_percent,na.rm=TRUE)
```

## 3.2 Coverage-breadth summary

```{r summarise-coverage-breadth}
breadth_summary <- breadth_long %>%
  group_by(stage,coverage_threshold) %>%
  summarise(
    samples=n(),
    median=median(breadth_percent,na.rm=TRUE),
    lower_IQR=quantile(breadth_percent,0.25,na.rm=TRUE),
    upper_IQR=quantile(breadth_percent,0.75,na.rm=TRUE),
    minimum=min(breadth_percent,na.rm=TRUE),
    maximum=max(breadth_percent,na.rm=TRUE),
    .groups="drop")
breadth_summary %>%
  mutate(across(c(median,lower_IQR,upper_IQR,minimum,maximum),~round(.x,1)))
```

## 3.3 Coverage-breadth plot

```{r plot-coverage-breadth}
breadth_colours<-c("Raw reads"="grey35","Duplex consensus"="darkcyan")

figure_3<-ggplot(breadth_summary,aes(x=coverage_threshold,y=median,colour=stage,fill=stage,group=stage))+
  geom_ribbon(aes(ymin=lower_IQR,ymax=upper_IQR),alpha=0.18,colour=NA)+
  geom_line(linewidth=1)+
  geom_point(size=2.5)+
  scale_colour_manual(values=breadth_colours)+
  scale_fill_manual(values=breadth_colours)+
  scale_x_log10(breaks=c(10,20,50,100,250,500,1000),labels=scales::label_comma())+
  scale_y_continuous(limits=c(0,100),breaks=seq(0,100,20),labels=function(x) paste0(x,"%"))+
  labs(x="Coverage threshold (×)",y="Target bases covered",colour="",fill="")+
  theme_classic(base_size=14)+
  theme(legend.position="bottom")

figure_3
ggsave(file.path(figure_path,"Figure3_coverage_breadth.pdf"),plot=figure_3,width=7,height=5)

ggsave(file.path(figure_path,"Figure3_coverage_breadth.png"),plot=figure_3,width=7,height=5,dpi=600)
```

# 4 Independent duplex-molecule yield

## 4.1 Load and prepare data

```{r load-molecule-counts}
molecule_file<-file.path(path,"independent_duplex_molecules.tsv")

molecule_data<-read.delim(molecule_file,sep="\t",header=TRUE,colClasses="character")

molecule_data<-molecule_data %>%
  filter(patient!="1452") %>%
  mutate(
    independent_duplex_molecules=
      as.numeric(independent_duplex_molecules),
    molecules_thousands=
      independent_duplex_molecules/1000,
    sample_group="Patient samples")

dim(molecule_data)

molecule_data %>%
  count(sample_group)
```

## 4.2 Molecule-yield summary

```{r summarise-molecule-yield}
molecule_summary<-molecule_data %>%
  summarise(
    samples=n(),
    median=median(molecules_thousands,na.rm=TRUE),
    Q1=quantile(molecules_thousands,0.25,na.rm=TRUE),
    Q3=quantile(molecules_thousands,0.75,na.rm=TRUE),
    minimum=min(molecules_thousands,na.rm=TRUE),
    maximum=max(
      molecules_thousands,
      na.rm=TRUE))

molecule_summary
```

## 4.3 Molecule-yield plot

```{r plot-molecule-yield}
figure_molecules<-ggplot(molecule_data,aes(x=sample_group,y=molecules_thousands))+
  geom_boxplot(width=0.4,outlier.shape=NA,fill="darkcyan",alpha=0.7)+
  geom_jitter(width=0.08,size=2,alpha=0.7)+
  scale_y_continuous(labels=scales::label_comma())+
  labs(x="",y="Mapped duplex-consensus molecules per sample (×10³)")+
  theme_classic(base_size=14)

figure_molecules
```

## 4.4 Save molecule-yield plot

```{r save-molecule-yield}
figure_path<-file.path(path,"figures")

dir.create(figure_path,showWarnings=FALSE)

ggsave(file.path(figure_path,"independent_duplex_molecules.pdf"),plot=figure_molecules,width=5.5,height=5)

ggsave(file.path(figure_path,"independent_duplex_molecules.png"),plot=figure_molecules,width=5.5,height=5,dpi=600)
```

# 5 Fold-80 coverage uniformity

## 5.1 Prepare data

```{r prepare-fold80}
uniformity_data<-hsmetrics %>%
  mutate(
    fold_80_base_penalty=
      as.numeric(fold_80_base_penalty),
    stage=recode(
      as.character(stage),
      "Raw"="Raw reads",
      "Duplex"="Duplex consensus",
      "Raw reads"="Raw reads",
      "Duplex consensus"="Duplex consensus"),
    stage=factor(
      stage,
      levels=c(
        "Raw reads",
        "Duplex consensus")))

uniformity_data %>%
  count(stage)
```

## 5.2 Uniformity summary

```{r summarise-fold80}
uniformity_summary<-uniformity_data %>%
  group_by(stage) %>%
  summarise(
    samples=sum(!is.na(fold_80_base_penalty)),
    median=median(fold_80_base_penalty,na.rm=TRUE),
    Q1=quantile(fold_80_base_penalty,0.25,na.rm=TRUE),
    Q3=quantile(fold_80_base_penalty,0.75,na.rm=TRUE),
    minimum=min(fold_80_base_penalty,na.rm=TRUE),
    maximum=max(fold_80_base_penalty,na.rm=TRUE),
    .groups="drop")

uniformity_summary
```

## 5.3 Pair samples

```{r pair-fold80}
uniformity_paired<-uniformity_data %>%
  select(sample,stage,fold_80_base_penalty) %>%
  mutate(stage=recode(as.character(stage),"Raw reads"="raw","Duplex consensus"="duplex")) %>%
  pivot_wider(names_from=stage,values_from=fold_80_base_penalty)

uniformity_change<-uniformity_paired %>%
  summarise(
    samples=n(),
    improved=sum(duplex<raw,na.rm=TRUE),
    unchanged=sum(duplex==raw,na.rm=TRUE),
    worsened=sum(
      duplex>raw,
      na.rm=TRUE))

uniformity_change
```

## 5.4 Fold-80 plot

```{r plot-fold80}
figure_uniformity<-ggplot(uniformity_paired,aes(x=raw,y=duplex))+
  geom_abline(slope=1,intercept=0,linetype="dashed",colour="grey50")+
  geom_point(colour="darkcyan",size=2.5,alpha=0.7)+
  coord_equal(xlim=c(1,1.6),ylim=c(1,1.6))+
  labs(x="Raw-read Fold-80 base penalty",y="Duplex-consensus Fold-80 base penalty")+
  theme_classic(base_size=14)

figure_uniformity
```

## 5.5 Save Fold-80 plot

```{r save-fold80}
ggsave(file.path(figure_path,"fold80_raw_vs_duplex.pdf"),plot=figure_uniformity,width=5.5,height=5)

ggsave(file.path(figure_path,"fold80_raw_vs_duplex.png"),plot=figure_uniformity,width=5.5,height=5,dpi=600)
```

# 6 Duplex-molecule recovery

## 6.1 Prepare data

```{r prepare-molecule-recovery}
recovery_data<-molecule_data %>%
  select(sample,patient,endpoint,independent_duplex_molecules) %>%
  inner_join(qc_yield %>% select(sample,read_pairs),by="sample") %>%
  mutate(
    read_pairs=as.numeric(read_pairs),
    independent_duplex_molecules=
      as.numeric(independent_duplex_molecules),
    read_pairs_million=
      read_pairs/1000000,
    molecules_thousands=
      independent_duplex_molecules/1000,
    recovery_percent=
      100*independent_duplex_molecules/read_pairs)

dim(recovery_data)

recovery_data %>%
  summarise(samples=n(),missing_yield=sum(is.na(read_pairs)),missing_molecules=sum(is.na(independent_duplex_molecules)))
```

## 6.2 Recovery summary

```{r summarise-molecule-recovery}
recovery_summary<-recovery_data %>%
  summarise(
    samples=n(),
    median=median(recovery_percent,na.rm=TRUE),
    Q1=quantile(recovery_percent,0.25,na.rm=TRUE),
    Q3=quantile(recovery_percent,0.75,na.rm=TRUE),
    minimum=min(recovery_percent,na.rm=TRUE),
    maximum=max(
      recovery_percent,
      na.rm=TRUE))

recovery_summary
```

## 6.3 Test sequencing and molecule yield

```{r test-molecule-recovery}
recovery_correlation<-cor.test(
  recovery_data$read_pairs_million,
  recovery_data$molecules_thousands,
  method="spearman",
  exact=FALSE)

recovery_correlation_label<-paste0(
  "No monotonic association: Spearman rho=",
  sprintf("%.2f",unname(recovery_correlation$estimate)),
  "; p=",format.pval(recovery_correlation$p.value,digits=3),
  "; n=",nrow(recovery_data))

recovery_correlation
```

## 6.4 Recovery plot

```{r plot-molecule-recovery}
figure_recovery<-ggplot(recovery_data,aes(x=read_pairs_million,y=molecules_thousands))+
  geom_point(shape=21,fill=unname(unified_colours["Patient samples"]),colour="black",size=2.2,stroke=0.35,alpha=0.8)+
  labs(
    x="Original sequencing yield (million read pairs)",
    y="Mapped duplex-consensus molecules (×10³)",
    subtitle=recovery_correlation_label)+
  theme_classic(base_size=14)

figure_recovery
```

## 6.5 Save recovery plot

```{r save-molecule-recovery}
ggsave(file.path(figure_path,"duplex_recovery_vs_sequencing_yield.pdf"),plot=figure_recovery,width=5.5,height=5)

ggsave(file.path(figure_path,"duplex_recovery_vs_sequencing_yield.png"),plot=figure_recovery,width=5.5,height=5,dpi=600)
```

## 6.6 Duplex family-size distribution

```{r duplex-family-size}
  family_size_file<-file.path(path,"duplex_family_sizes.tsv")

  patient_family_sizes<-read_tsv(family_size_file,show_col_types=FALSE) %>%
    mutate(
      patient=as.character(patient),
      endpoint=toupper(as.character(endpoint)),
      ab_size=as.numeric(ab_size),
      ba_size=as.numeric(ba_size),
      count=as.numeric(count)) %>%
    filter(patient!="1452",endpoint%in%c("BL","END"))

  family_size_valid<-patient_family_sizes %>%
    filter(is.finite(ab_size),is.finite(ba_size),is.finite(count),ab_size>=1,ba_size>=0,count>0)

  #Calculate family-support thresholds
  family_support_by_sample<-family_size_valid %>%
    group_by(sample,patient,endpoint) %>%
    summarise(
      total_families=sum(count),
      ba_gt0_families=sum(count[ba_size>0]),
      ge2_families=sum(count[ab_size>=2 & ba_size>=2]),
      ge3_families=sum(count[ab_size>=3 & ba_size>=3]),
      ge5_families=sum(count[ab_size>=5 & ba_size>=5]),
      .groups="drop")

  family_threshold_reference<-tribble(
    ~metric,            ~threshold,                 ~definition,             ~threshold_order,
    "ba_gt0_families", "BA > 0 (duplex yield)",   "AB >=1 and BA >=1",                  1,
    "ge2_families",    ">=2+2",                   "AB >=2 and BA >=2",                  2,
    "ge3_families",    ">=3+3",                   "AB >=3 and BA >=3",                  3,
    "ge5_families",    ">=5+5",                   "AB >=5 and BA >=5",                  4)

  family_support_long<-family_support_by_sample %>%
    select(sample,total_families,ba_gt0_families,ge2_families,ge3_families,ge5_families) %>%
    pivot_longer(
      cols=c(ba_gt0_families,ge2_families,ge3_families,ge5_families),
      names_to="metric",
      values_to="families") %>%
    left_join(family_support_by_sample %>% select(sample,ba_gt0_families),by="sample") %>%
    mutate(sample_percent=100*families/total_families)

  family_support_summary<-family_support_long %>%
    group_by(metric) %>%
    summarise(
      samples=n_distinct(sample),
      families=sum(families),
      total_families=sum(total_families),
      pooled_percent=100*families/total_families,
      median_percent=median(sample_percent),
      Q1=quantile(sample_percent,0.25),
      Q3=quantile(sample_percent,0.75),
      minimum=min(sample_percent),
      maximum=max(sample_percent),
      among_duplex_percent=100*families/sum(ba_gt0_families),
      .groups="drop") %>%
    left_join(family_threshold_reference,by="metric") %>%
    arrange(threshold_order)

  family_support_medians<-setNames(family_support_summary$median_percent,family_support_summary$metric)

  #Check the duplex-yield calculation
  duplex_yield_validation<-read_tsv(file.path(path,"duplex_yield_100pct.tsv"),show_col_types=FALSE) %>%
    transmute(
      sample=as.character(sample),
      patient=as.character(patient),
      reported_duplex_yield_percent=
        100*as.numeric(ds_fraction_duplexes)) %>%
    filter(patient!="1452") %>%
    inner_join(
      family_support_by_sample %>%
        transmute(sample,calculated_duplex_yield_percent= 100*ba_gt0_families/total_families),
      by="sample") %>%
    mutate(absolute_difference=abs(calculated_duplex_yield_percent- reported_duplex_yield_percent))

  stopifnot(
    nrow(duplex_yield_validation)==
      nrow(family_support_by_sample),
    max(duplex_yield_validation$absolute_difference)<0.001)

  duplex_yield_summary<-duplex_yield_validation %>%
    summarise(
      samples=n(),
      median=median(reported_duplex_yield_percent),
      Q1=quantile(reported_duplex_yield_percent,0.25),
      Q3=quantile(reported_duplex_yield_percent,0.75),
      minimum=min(reported_duplex_yield_percent),
      maximum=max(reported_duplex_yield_percent),
      maximum_validation_difference=max(absolute_difference))

  family_support_table<-family_support_summary %>%
    transmute(
      threshold,
      definition,
      pooled_families=paste0(label_comma()(families)," / ",label_comma()(total_families)),
      pooled_percent,
      median_IQR=paste0(sprintf("%.1f",median_percent)," (",sprintf("%.1f",Q1),"-",sprintf("%.1f",Q3),")"),
      among_duplex_percent)

  family_support_gt<-family_support_table %>%
    gt() %>%
    cols_label(
      threshold="Family support threshold",
      definition="Definition",
      pooled_families="Pooled families, n/N",
      pooled_percent="Pooled families",
      median_IQR="Median per sample (IQR), %",
      among_duplex_percent="Among duplex families") %>%
    fmt_number(columns=c(pooled_percent,among_duplex_percent),decimals=1,suffix="%") %>%
    cols_align("left",columns=c(threshold,definition)) %>%
    cols_align("center",columns=-c(threshold,definition)) %>%
    tab_header(
      title=md("**UMI-family support and observed duplex yield**"),
      subtitle=paste0(
        nrow(family_support_by_sample),
        " patient samples; thresholds calculated from unbinned family counts")) %>%
    tab_style(style=cell_text(weight="bold"),locations=cells_column_labels()) %>%
    tab_source_note(md(paste0(
      "*BA > 0 is the observed duplex yield: ds_duplexes / ds_families. ",
      "It matched the fgbio value in every sample (maximum absolute difference ",
      sprintf("%.4f",duplex_yield_summary$maximum_validation_difference),
      " percentage points).*"))) %>%
    tab_source_note(md(
      "*Pooled percentages weight samples by their number of UMI families; the median and IQR give every sample equal weight.*")) %>%
    tab_options(
      table.font.names="Arial",
      table.font.size=11,
      heading.title.font.size=11,
      heading.subtitle.font.size=10,
      heading.align="left",
      data_row.padding=2,
      column_labels.padding=4,
      source_notes.font.size=9,
      row.striping.background_color="white")

  family_size_limit<-10

  family_size_sample<-family_size_valid %>%
    mutate(ab_plot=pmin(ab_size,family_size_limit),ba_plot=pmin(ba_size,family_size_limit)) %>%
    group_by(sample,ab_plot,ba_plot) %>%
    summarise(families=sum(count),.groups="drop") %>%
    complete(sample,ab_plot=1:family_size_limit,ba_plot=0:family_size_limit,fill=list(families=0)) %>%
    filter(ba_plot<=ab_plot) %>%
    group_by(sample) %>%
    mutate(total_families=sum(families),family_percent=100*families/total_families) %>%
    ungroup()

  family_size_summary<-family_size_sample %>%
    group_by(ab_plot,ba_plot) %>%
    summarise(
      samples=n_distinct(sample),
      median_percent=median(family_percent,na.rm=TRUE),
      Q1=quantile(family_percent,0.25,na.rm=TRUE),
      Q3=quantile(family_percent,0.75,na.rm=TRUE),
      .groups="drop")

  duplex_family_size_levels<-c("1+1","2+2","3+3","4+4",">=5+5")

  duplex_family_size_counts<-family_size_valid %>%
    filter(ba_size>0) %>%
    mutate(
      family_size_category=case_when(
        ba_size==1 ~ "1+1",
        ba_size==2 ~ "2+2",
        ba_size==3 ~ "3+3",
        ba_size==4 ~ "4+4",
        ba_size>=5 ~ ">=5+5")) %>%
    group_by(sample,family_size_category) %>%
    summarise(families=sum(count),.groups="drop")

  duplex_family_totals<-family_size_valid %>%
    filter(ba_size>0) %>%
    group_by(sample) %>%
    summarise(duplex_families=sum(count),.groups="drop")

  duplex_family_size_by_sample<-expand_grid(
      sample=duplex_family_totals$sample,
      family_size_category=duplex_family_size_levels) %>%
    left_join(duplex_family_size_counts,by=c("sample","family_size_category")) %>%
    left_join(duplex_family_totals,by="sample") %>%
    mutate(
      families=replace_na(families,0),
      family_percent=100*families/duplex_families,
      family_size_category=factor(
        family_size_category,
        levels=duplex_family_size_levels))

  duplex_family_size_check<-duplex_family_size_by_sample %>%
    group_by(sample) %>%
    summarise(total_percent=sum(family_percent),.groups="drop")

  stopifnot(all(abs(duplex_family_size_check$total_percent-100)<1e-8))

  duplex_family_size_summary<-duplex_family_size_by_sample %>%
    group_by(family_size_category) %>%
    summarise(
      samples=n_distinct(sample),
      median_percent=median(family_percent,na.rm=TRUE),
      Q1=quantile(family_percent,0.25,na.rm=TRUE),
      Q3=quantile(family_percent,0.75,na.rm=TRUE),
      .groups="drop")

  figure_duplex_family_size<-ggplot(duplex_family_size_summary,aes(x=family_size_category,y=median_percent))+
    geom_col(width=0.68,fill=unname(unified_colours["Patient samples"]),colour="black",linewidth=0.35,alpha=0.80)+
    geom_errorbar(aes(ymin=Q1,ymax=Q3),width=0.16,linewidth=0.5)+
    geom_text(aes(y=Q3,label=paste0(sprintf("%.1f",median_percent),"%")),vjust=-0.45,size=3.2)+
    scale_y_continuous(labels=label_percent(scale=1),expand=expansion(mult=c(0,0.14)))+
    labs(
      x="Minimum reads in each strand family",
      y="Duplex families per sample (%)",
      subtitle="Bars show medians; whiskers show the interquartile range",
      caption=paste0(
        "Categories use BA, the smaller strand family; ",
        "the larger AB family may contain additional reads."))+
    theme_unified()+
    theme(
      panel.grid=element_blank(),
      plot.subtitle=element_text(size=9),
      plot.caption=element_text(size=8,hjust=0,margin=margin(t=5)))

  figure_duplex_family_size
  family_support_gt
```

# 7 Duplex fraction and sequencing saturation

## 7.1 Load full-depth duplex data

```{r load-duplex-fractions}
duplex_file<-file.path(path,"duplex_yield_100pct.tsv")

if(!file.exists(duplex_file)){
  stop("Missing file: duplex_yield_100pct.tsv")
}

duplex_data<-read.delim(duplex_file,sep="\t",header=TRUE,colClasses="character")

duplex_data<-duplex_data %>%
  filter(patient!="1452") %>%
  mutate(
    read_pairs=as.numeric(read_pairs),
    ds_families=as.numeric(ds_families),
    ds_duplexes=as.numeric(ds_duplexes),
    ds_fraction_duplexes=
      as.numeric(ds_fraction_duplexes),
    ds_fraction_duplexes_ideal=
      as.numeric(ds_fraction_duplexes_ideal))

dim(duplex_data)
```

## 7.2 Validate duplex fractions

```{r validate-duplex-fractions}
range(duplex_data$ds_fraction_duplexes,na.rm=TRUE)

range(duplex_data$ds_fraction_duplexes_ideal,na.rm=TRUE)

if(any(
  duplex_data$ds_fraction_duplexes<0 |
  duplex_data$ds_fraction_duplexes>1,
  na.rm=TRUE)){
  stop("ds_fraction_duplexes is not a proportion between 0 and 1")
}

if(any(
  duplex_data$ds_fraction_duplexes_ideal<0 |
  duplex_data$ds_fraction_duplexes_ideal>1,
  na.rm=TRUE)){
  stop("ds_fraction_duplexes_ideal is not a proportion between 0 and 1")
}
```

```{r convert-duplex-percent}
duplex_data<-duplex_data %>%
  mutate(
    duplex_fraction=
      100*ds_fraction_duplexes,
    ideal_duplex_fraction=
      100*ds_fraction_duplexes_ideal,
    duplex_fraction_gap=
      ideal_duplex_fraction-duplex_fraction)
```

## 7.3 Full-depth duplex summary

```{r summarise-duplex-fractions}
duplex_summary<-duplex_data %>%
  summarise(
    samples=n(),
    observed_median=median(duplex_fraction,na.rm=TRUE),
    observed_Q1=quantile(duplex_fraction,0.25,na.rm=TRUE),
    observed_Q3=quantile(duplex_fraction,0.75,na.rm=TRUE),
    observed_minimum=min(duplex_fraction,na.rm=TRUE),
    observed_maximum=max(duplex_fraction,na.rm=TRUE),
    ideal_median=median(ideal_duplex_fraction,na.rm=TRUE),
    ideal_Q1=quantile(ideal_duplex_fraction,0.25,na.rm=TRUE),
    ideal_Q3=quantile(ideal_duplex_fraction,0.75,na.rm=TRUE),
    median_gap=median(
      duplex_fraction_gap,
      na.rm=TRUE))

duplex_summary
```

## 7.4 Observed and ideal duplex fractions

```{r plot-duplex-fractions}
figure_duplex_paired<-ggplot(duplex_data,aes(x=duplex_fraction,y=ideal_duplex_fraction))+
  geom_abline(slope=1,intercept=0,linetype="dashed",colour="grey50")+
  geom_point(colour="darkcyan",size=2.5,alpha=0.7)+
  coord_equal(xlim=c(50,100),ylim=c(50,100))+
  labs(x="Observed duplex fraction (%)",y="Ideal duplex fraction (%)")+
  theme_classic(base_size=14)

figure_duplex_paired
```

## 7.5 Save duplex-fraction plot

```{r save-duplex-fractions}
ggsave(file.path(figure_path,"observed_vs_ideal_duplex_fraction.pdf"),plot=figure_duplex_paired,width=5.5,height=5)

ggsave(
  file.path(figure_path,"observed_vs_ideal_duplex_fraction.png"),
  plot=figure_duplex_paired,
  width=5.5,
  height=5,
  dpi=600)
```

## 7.6 Load saturation data

```{r load-saturation-data}
saturation_file<-file.path(path,"duplex_yield_all_fractions.tsv")

if(!file.exists(saturation_file)){
  stop("Missing file: duplex_yield_all_fractions.tsv")
}

saturation_data<-read.delim(saturation_file,sep="\t",header=TRUE,colClasses="character")

saturation_data<-saturation_data %>%
  filter(patient!="1452") %>%
  mutate(
    fraction=as.numeric(fraction),
    read_pairs=as.numeric(read_pairs),
    ds_families=as.numeric(ds_families),
    ds_duplexes=as.numeric(ds_duplexes),
    ds_fraction_duplexes=
      as.numeric(ds_fraction_duplexes),
    ds_fraction_duplexes_ideal=
      as.numeric(ds_fraction_duplexes_ideal))

dim(saturation_data)
```

## 7.7 Validate saturation proportions

```{r validate-saturation-proportions}
range(saturation_data$fraction,na.rm=TRUE)

range(saturation_data$ds_fraction_duplexes,na.rm=TRUE)

range(saturation_data$ds_fraction_duplexes_ideal,na.rm=TRUE)

saturation_values<-c(
  saturation_data$fraction,
  saturation_data$ds_fraction_duplexes,
  saturation_data$ds_fraction_duplexes_ideal)

if(any(
  saturation_values<0 |
  saturation_values>1,
  na.rm=TRUE)){
  stop("One or more saturation variables are not proportions between 0 and 1")
}
```

```{r convert-saturation-percent}
saturation_data<-saturation_data %>%
  mutate(
    sequencing_fraction=
      100*fraction,
    observed_duplex_fraction=
      100*ds_fraction_duplexes,
    ideal_duplex_fraction=
      100*ds_fraction_duplexes_ideal)
```

```{r check-saturation-samples}
saturation_data %>%
  count(sequencing_fraction)
```

## 7.8 Saturation summary

```{r summarise-saturation}
saturation_summary<-saturation_data %>%
  group_by(sequencing_fraction) %>%
  summarise(
    samples=n(),
    observed_median=median(observed_duplex_fraction,na.rm=TRUE),
    observed_Q1=quantile(observed_duplex_fraction,0.25,na.rm=TRUE),
    observed_Q3=quantile(observed_duplex_fraction,0.75,na.rm=TRUE),
    ideal_median=median(ideal_duplex_fraction,na.rm=TRUE),
    ideal_Q1=quantile(ideal_duplex_fraction,0.25,na.rm=TRUE),
    ideal_Q3=quantile(ideal_duplex_fraction,0.75,na.rm=TRUE),
    .groups="drop")

saturation_summary
```

## 7.9 Saturation curve

```{r plot-saturation}
figure_saturation<-ggplot(saturation_summary,aes(x=sequencing_fraction))+
  geom_ribbon(aes(ymin=observed_Q1,ymax=observed_Q3),fill="darkcyan",alpha=0.18)+
  geom_ribbon(aes(ymin=ideal_Q1,ymax=ideal_Q3),fill="grey60",alpha=0.12)+
  geom_line(aes(y=observed_median,colour="Observed duplex fraction"),linewidth=1)+
  geom_line(aes(y=ideal_median,colour="Ideal duplex fraction"),linewidth=1,linetype="dashed")+
  scale_colour_manual(values=c("Observed duplex fraction"="darkcyan","Ideal duplex fraction"="grey35"))+
  scale_x_continuous(breaks=seq(0,100,20),limits=c(0,100))+
  scale_y_continuous(breaks=seq(0,100,20),limits=c(0,100),labels=scales::label_percent(scale=1))+
  labs(x="Sequencing data used (%)",y="Duplex fraction",colour="")+
  theme_classic(base_size=14)+
  theme(legend.position="bottom")

figure_saturation
```

## 7.10 Save saturation curve

```{r save-saturation}
ggsave(file.path(figure_path,"duplex_saturation_curve.pdf"),plot=figure_saturation,width=6.5,height=5)

ggsave(file.path(figure_path,"duplex_saturation_curve.png"),plot=figure_saturation,width=6.5,height=5,dpi=600)
```

# 8 HsMetrics technical summary

## 8.1 Load data

```{r load-full-hsmetrics}
full_hsmetrics_file<-file.path(path,"full_hsmetrics.tsv")

file.exists(full_hsmetrics_file)

full_hsmetrics<-read.delim(full_hsmetrics_file,sep="\t",header=TRUE,colClasses="character",check.names=FALSE)

dim(full_hsmetrics)

full_hsmetrics %>%
  count(endpoint)
```

## 8.2 Prepare variables

```{r prepare-full-hsmetrics}
full_hsmetrics<-full_hsmetrics %>%
  filter(patient!="1452") %>%
  mutate(
    sample_group=if_else(endpoint=="HC","Healthy controls","Patient samples"),
    mean_target_coverage=
      as.numeric(MEAN_TARGET_COVERAGE),
    median_target_coverage=
      as.numeric(MEDIAN_TARGET_COVERAGE),
    selected_bases_percent=
      100*as.numeric(PCT_SELECTED_BASES),
    usable_bases_on_target_percent=
      100*as.numeric(PCT_USABLE_BASES_ON_TARGET),
    fold_80_base_penalty=
      as.numeric(FOLD_80_BASE_PENALTY),
    zero_coverage_targets_percent=
      100*as.numeric(ZERO_CVG_TARGETS_PCT),
    target_bases_1000x_percent=
      100*as.numeric(PCT_TARGET_BASES_1000X),
    target_bases_2500x_percent=
      100*as.numeric(PCT_TARGET_BASES_2500X),
    target_bases_5000x_percent=
      100*as.numeric(PCT_TARGET_BASES_5000X))

full_hsmetrics %>%
  count(sample_group)
```

## 8.3 HsMetrics summary

```{r summarise-full-hsmetrics}
full_hsmetrics_summary<-full_hsmetrics %>%
  group_by(sample_group) %>%
  summarise(
    samples=n(),
    median_mean_coverage=median(mean_target_coverage,na.rm=TRUE),
    median_selected_bases_percent=median(selected_bases_percent,na.rm=TRUE),
    median_usable_on_target_percent=median(usable_bases_on_target_percent,na.rm=TRUE),
    median_fold_80=median(fold_80_base_penalty,na.rm=TRUE),
    median_zero_coverage_targets_percent=median(zero_coverage_targets_percent,na.rm=TRUE),
    median_target_bases_1000x_percent=median(target_bases_1000x_percent,na.rm=TRUE),
    median_target_bases_2500x_percent=median(target_bases_2500x_percent,na.rm=TRUE),
    median_target_bases_5000x_percent=median(target_bases_5000x_percent,na.rm=TRUE),
    .groups="drop")

full_hsmetrics_summary_display<-full_hsmetrics_summary %>%
  mutate(across(where(is.numeric),~round(.x,2)))

full_hsmetrics_summary_display
```

```{r format-hsmetrics-table}
hsmetrics_table<-full_hsmetrics_summary_display %>%
  transmute(
    Group=sample_group,
    N=samples,
    `Mean depth (×)`=median_mean_coverage,
    `Selected bases (%)`=median_selected_bases_percent,
    `Usable on-target (%)`=median_usable_on_target_percent,
    `Fold-80`=median_fold_80,
    `Zero coverage (%)`=median_zero_coverage_targets_percent,
    `≥1000× (%)`=median_target_bases_1000x_percent,
    `≥2500× (%)`=median_target_bases_2500x_percent,
    `≥5000× (%)`=median_target_bases_5000x_percent)

hsmetrics_gt<-hsmetrics_table %>%
  gt() %>%
  tab_header(title="Picard HsMetrics summary for final duplex-consensus BAMs") %>%
  tab_options(table.font.size=px(11))

hsmetrics_gt
```

# 9 Per-target and per-gene coverage

## 9.1 Load per-target coverage

```{r load-per-target-coverage}
per_target_file<-file.path(path,"per_target_coverage.tsv")

file.exists(per_target_file)

per_target<-read.delim(per_target_file,sep="\t",header=TRUE,colClasses="character",check.names=FALSE)

dim(per_target)

names(per_target)
```

```{r check-target-counts}
per_target %>%
  count(sample) %>%
  count(n)
```

## 9.2 Prepare target data

```{r prepare-target-coverage}
target_data<-per_target %>%
  filter(patient!="1452") %>%
  mutate(
    sample_group=if_else(endpoint=="HC","Healthy controls","Patient samples"),
    length=as.numeric(length),
    mean_coverage=as.numeric(mean_coverage),
    min_coverage=as.numeric(min_coverage),
    max_coverage=as.numeric(max_coverage))

target_data %>%
  count(sample_group)
```

## 9.3 Calculate gene-level coverage

```{r calculate-gene-coverage}
gene_coverage<-target_data %>%
  group_by(sample,patient,endpoint,sample_group,name) %>%
  summarise(
    target_intervals=n(),
    target_bases=sum(length,na.rm=TRUE),
    mean_gene_coverage=weighted.mean(mean_coverage,w=length,na.rm=TRUE),
    minimum_base_coverage=min(min_coverage,na.rm=TRUE),
    .groups="drop") %>%
  rename(gene=name)

gene_coverage %>%
  count(sample) %>%
  count(n)
```

## 9.4 Gene-coverage summary

```{r summarise-gene-coverage}
gene_coverage_summary<-gene_coverage %>%
  group_by(sample_group,gene) %>%
  summarise(
    samples=n(),
    median=median(mean_gene_coverage,na.rm=TRUE),
    Q1=quantile(mean_gene_coverage,0.25,na.rm=TRUE),
    Q3=quantile(mean_gene_coverage,0.75,na.rm=TRUE),
    minimum=min(mean_gene_coverage,na.rm=TRUE),
    maximum=max(mean_gene_coverage,na.rm=TRUE),
    .groups="drop")

gene_coverage_summary
```

## 9.5 Gene-coverage plot

```{r plot-gene-coverage}
gene_coverage<-gene_coverage %>% mutate(gene=factor(gene,levels=c("BAP1","SF3B1","EIF1AX","MBD4","GNAQ","GNA11","PLCB4","CYSLTR2")))

gene_colours<-c(
  "Healthy controls"=unname(unified_colours["Healthy controls"]),
  "Patient samples"=unname(unified_colours["Patient samples"]))

figure_gene_coverage<-ggplot(gene_coverage,aes(x=gene,y=mean_gene_coverage,fill=sample_group,colour=sample_group))+
  geom_boxplot(position=position_dodge(width=0.75),width=0.6,outlier.shape=NA,alpha=0.6)+
  geom_point(position=position_jitterdodge(jitter.width=0.08,dodge.width=0.75),size=1.5,alpha=0.6)+
  scale_fill_manual(values=gene_colours)+
  scale_colour_manual(values=gene_colours)+
  scale_y_continuous(labels=scales::label_comma())+
  labs(x="Targeted gene",y="Mean duplex-consensus coverage (×)") +
theme_unified() +
theme(axis.text.x=element_text(angle=45,hjust=1),legend.position="bottom")

figure_gene_coverage <- standardize_plot(figure_gene_coverage)
figure_gene_coverage$layers[[2]]$aes_params$colour <- "black"
figure_gene_coverage
```

# 10 Base-level panel coverage

## 10.1 Load coverage summaries

```{r load-base-coverage}
base_sample_file<-file.path(path,"base_depth_sample_summary.tsv")
base_gene_file<-file.path(path,"base_depth_gene_summary.tsv")

file.exists(base_sample_file)
file.exists(base_gene_file)

base_sample<-read.delim(base_sample_file,sep="\t",header=TRUE,check.names=FALSE)
base_gene<-read.delim(base_gene_file,sep="\t",header=TRUE,check.names=FALSE)

dim(base_sample)
dim(base_gene)

names(base_sample)
names(base_gene)
```

```{r prepare-base-coverage}
base_sample<-base_sample %>%
  mutate(sample_group=if_else(endpoint=="HC","Healthy controls","Patient samples"))

base_gene<-base_gene %>%
  mutate(sample_group=if_else(endpoint=="HC","Healthy controls","Patient samples"))

base_sample %>% count(sample_group)
base_gene %>% count(sample_group)
```

## 10.2 Base-coverage summary by sample

```{r summarise-base-coverage}
base_depth_summary<-base_sample %>%
  group_by(sample_group) %>%
  summarise(
    samples=n(),
    median_mean_depth=median(mean_depth,na.rm=TRUE),
    median_minimum_depth=median(minimum_depth,na.rm=TRUE),
    median_zero_percent=median(zero_percent,na.rm=TRUE),
    median_250x=median(pct_250x,na.rm=TRUE),
    median_500x=median(pct_500x,na.rm=TRUE),
    median_1000x=median(pct_1000x,na.rm=TRUE),
    median_2000x=median(pct_2000x,na.rm=TRUE),
    median_3000x=median(pct_3000x,na.rm=TRUE),
    .groups="drop")

base_depth_summary
```

## 10.3 Prepare coverage thresholds

```{r prepare-coverage-thresholds}
base_threshold_data<-base_sample %>%
  select(sample,sample_group,pct_250x,pct_500x,pct_1000x,pct_2000x,pct_3000x) %>%
  pivot_longer(cols=starts_with("pct_"),names_to="threshold",values_to="percent_covered") %>%
  mutate(
    threshold=recode(
      threshold,
      "pct_250x"="250×",
      "pct_500x"="500×",
      "pct_1000x"="1,000×",
      "pct_2000x"="2,000×",
      "pct_3000x"="3,000×"),
    threshold=factor(threshold,levels=c("250×","500×","1,000×","2,000×","3,000×")))

base_threshold_data %>% count(sample_group,threshold)
```

```{r summarise-coverage-thresholds}
base_threshold_summary<-base_threshold_data %>%
  group_by(sample_group,threshold) %>%
  summarise(
    samples=n(),
    median=median(percent_covered,na.rm=TRUE),
    Q1=quantile(percent_covered,0.25,na.rm=TRUE),
    Q3=quantile(percent_covered,0.75,na.rm=TRUE),
    .groups="drop")

base_threshold_summary
```

## 10.4 Plot coverage thresholds

```{r plot-coverage-thresholds}
group_colours<-c(
  "Healthy controls"=unname(unified_colours["Healthy controls"]),
  "Patient samples"=unname(unified_colours["Patient samples"]))

figure_base_depth<-ggplot(
  base_threshold_summary,
  aes(x=threshold,y=median,colour=sample_group,fill=sample_group,group=sample_group))+
  geom_ribbon(aes(ymin=Q1,ymax=Q3),colour=NA,alpha=0.15)+
  geom_line(linewidth=0.8)+
  geom_point(size=2)+
  scale_colour_manual(values=group_colours)+
  scale_fill_manual(values=group_colours)+
  scale_y_continuous(limits=c(0,100),breaks=seq(0,100,20),labels=scales::label_percent(scale=1))+
  labs(x="Coverage threshold",y="Target bases covered",colour="",fill="")+
  theme_unified()+
  theme(legend.position="bottom")

figure_base_depth
```

## 10.5 Target-interval coverage

### 10.5.1 Prepare interval data

```{r prepare-interval-data}
per_target_file<-file.path(path,"per_target_coverage.tsv")

per_target<-read.delim(per_target_file,sep="\t",header=TRUE,check.names=FALSE) %>%
  filter(patient!="1452") %>%
  mutate(
    start=as.numeric(start),
    end=as.numeric(end),
    mean_coverage=as.numeric(mean_coverage),
    endpoint=recode(endpoint,"BL"="Baseline","END"="Post-radiotherapy","HC"="Healthy controls"),
    endpoint=factor(
      endpoint,
      levels=c(
        "Baseline",
        "Post-radiotherapy",
        "Healthy controls")))

n_distinct(per_target$sample)
n_distinct(paste(per_target$chrom,per_target$start,per_target$end))
```

### 10.5.2 Create interval labels

```{r create-interval-labels}
target_key<-per_target %>%
  distinct(chrom,start,end,name) %>%
  arrange(name,chrom,start,end) %>%
  group_by(name) %>%
  mutate(interval_number=row_number(),interval=paste0(name,"_",interval_number)) %>%
  ungroup()

interval_data<-per_target %>%
  left_join(target_key,by=c("chrom","start","end","name"))

sample_order<-interval_data %>%
  distinct(sample,patient,endpoint) %>%
  mutate(patient_number=as.numeric(patient)) %>%
  arrange(endpoint,patient_number,sample) %>%
  pull(sample)

interval_order<-target_key %>%
  arrange(name,chrom,start,end) %>%
  pull(interval)

interval_data<-interval_data %>%
  mutate(sample=factor(sample,levels=sample_order),interval=factor(interval,levels=rev(interval_order)))
```

### 10.5.3 Interval coverage heatmap

```{r interval-level-coverage-heatmap, fig.width=12, fig.height=11.5, dpi=150}
patient_interval_heatmap_data<-interval_data %>%
  filter(endpoint%in%c("Baseline","Post-radiotherapy"))

figure_interval_heatmap<-ggplot(patient_interval_heatmap_data,aes(x=sample,y=interval,fill=mean_coverage))+
  geom_tile()+
  facet_grid(
    name~endpoint,
    scales="free",
    space="free_y",
    switch="y",
    labeller=labeller(
      endpoint=as_labeller(
        c(
          "Baseline"="Baseline",
          "Post-radiotherapy"="Post-radiotherapy"))))+
  scale_fill_viridis_c(trans=scales::pseudo_log_trans(sigma=1),labels=scales::label_comma(),name="Mean coverage\n(x)")+
  scale_y_discrete(labels=function(x) sub("^.*_", "", x))+
  labs(x="Samples",y="Capture interval")+
  theme_classic(base_size=10)+
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y=element_text(size=7),
    axis.title.y=element_text(margin=margin(r=10)),
    strip.background=element_blank(),
    strip.placement="outside",
    strip.text.x=element_text(size=11),
    strip.text.y.left=element_text(angle=0,face="bold",size=9,margin=margin(r=5)),
    panel.spacing.x=grid::unit(2,"mm"),
    panel.spacing.y=grid::unit(1,"mm"),
    plot.margin=margin(
      10,
      20,
      10,
      35))

figure_interval_heatmap
```

### 10.5.4 Save interval heatmap

```{r save-interval-heatmap}
ggsave(
  file.path(figure_path,"Supplementary_interval_coverage_heatmap.pdf"),
  plot=figure_interval_heatmap,
  width=12,
  height=11.5)

ggsave(
  file.path(figure_path,"Supplementary_interval_coverage_heatmap.png"),
  plot=figure_interval_heatmap,
  width=12,
  height=11.5,
  dpi=600)
```

### 10.5.5 Interval summary

```{r summarise-intervals}
interval_summary<-interval_data %>%
  group_by(name,interval,chrom,start,end) %>%
  summarise(
    samples=n(),
    median_coverage=median(mean_coverage,na.rm=TRUE),
    minimum_coverage=min(mean_coverage,na.rm=TRUE),
    maximum_coverage=max(mean_coverage,na.rm=TRUE),
    .groups="drop")

interval_summary
```

### 10.5.6 Patient interval summary

```{r summarise-patient-intervals}
patient_interval_summary<-interval_data %>%
  filter(endpoint!="Healthy controls") %>%
  group_by(name,interval,chrom,start,end) %>%
  summarise(
    samples=n(),
    median_coverage=median(mean_coverage,na.rm=TRUE),
    minimum_coverage=min(mean_coverage,na.rm=TRUE),
    maximum_coverage=max(mean_coverage,na.rm=TRUE),
    samples_below_500x=sum(mean_coverage<500,na.rm=TRUE),
    samples_below_1000x=sum(mean_coverage<1000,na.rm=TRUE),
    .groups="drop") %>%
  arrange(median_coverage)

patient_interval_summary
patient_interval_summary %>%
  slice_head(n=10)
patient_interval_check<-patient_interval_summary %>%
  summarise(
    intervals=n(),
    intervals_with_zero_minimum=
      sum(minimum_coverage==0),
    intervals_with_minimum_below_500x=
      sum(minimum_coverage<500),
    intervals_with_median_below_1000x=
      sum(median_coverage<1000))

patient_interval_check
```

# 11 Samples requiring QC review

## 11.1 Combine coverage and molecule data

```{r combine-patient-qc}
patient_qc<-base_sample %>%
  filter(endpoint %in% c("BL","END")) %>%
  left_join(molecule_data %>% select(sample,independent_duplex_molecules),by="sample") %>%
  mutate(independent_duplex_molecules=as.numeric(independent_duplex_molecules))

patient_qc %>%
  summarise(samples=n(),missing_molecules=sum(is.na(independent_duplex_molecules)))
```

## 11.2 Calculate the expected range

```{r calculate-qc-limits}
qc_limits<-patient_qc %>%
  summarise(
    lower_mean_depth=max(0,quantile(mean_depth,0.25,na.rm=TRUE)-1.5*IQR(mean_depth,na.rm=TRUE)),
    lower_500x=max(0,quantile(pct_500x,0.25,na.rm=TRUE)-1.5*IQR(pct_500x,na.rm=TRUE)),
    lower_molecules=max(0,quantile(independent_duplex_molecules,0.25,na.rm=TRUE)-1.5*IQR(independent_duplex_molecules,na.rm=TRUE)),
    upper_zero_percent=quantile(zero_percent,0.75,na.rm=TRUE)+1.5*IQR(zero_percent,na.rm=TRUE))

qc_limits
```

## 11.3 Flag samples for review

```{r flag-qc-samples}
patient_qc<-patient_qc %>%
  mutate(
    qc_status=if_else(
      is.na(independent_duplex_molecules) |
      mean_depth<qc_limits$lower_mean_depth |
      pct_500x<qc_limits$lower_500x |
      independent_duplex_molecules<qc_limits$lower_molecules |
      zero_percent>qc_limits$upper_zero_percent,
      "Review",
      "Within cohort range"))

patient_qc %>% count(qc_status)
```

## 11.4 Display samples for review

```{r display-qc-samples}
samples_for_review<-patient_qc %>%
  filter(qc_status=="Review") %>%
  select(sample,endpoint,mean_depth,independent_duplex_molecules,zero_percent,pct_500x,pct_1000x) %>%
  arrange(mean_depth)

samples_for_review
```

# 12 Baseline and post-radiotherapy QC balance

## 12.1 Combine timepoint measurements

```{r combine-timepoint-qc}
timepoint_qc<-base_sample %>%
  filter(endpoint %in% c("BL","END")) %>%
  select(sample,patient,endpoint,mean_depth,pct_1000x) %>%
  left_join(qc_yield %>% select(sample,read_pairs),by="sample") %>%
  left_join(molecule_data %>% select(sample,independent_duplex_molecules),by="sample") %>%
  mutate(
    read_pairs=as.numeric(read_pairs),
    independent_duplex_molecules=as.numeric(independent_duplex_molecules),
    read_pairs_million=read_pairs/1000000,
    molecules_thousands=independent_duplex_molecules/1000,
    timepoint=recode(endpoint,"BL"="Baseline","END"="Post-radiotherapy"),
    timepoint=factor(timepoint,levels=c("Baseline","Post-radiotherapy")))

timepoint_qc %>%
  summarise(
    samples=n(),
    patients=n_distinct(patient),
    missing_yield=sum(is.na(read_pairs)),
    missing_molecules=sum(is.na(independent_duplex_molecules)))
```

## 12.2 Confirm patient pairs

```{r check-timepoint-pairs}
timepoint_qc %>%
  count(patient) %>%
  count(n)
```

## 12.3 Summarise QC by timepoint

```{r summarise-timepoint-qc}
timepoint_summary<-timepoint_qc %>%
  group_by(timepoint) %>%
  summarise(
    samples=n(),
    median_read_pairs_million=median(read_pairs_million,na.rm=TRUE),
    median_molecules_thousands=median(molecules_thousands,na.rm=TRUE),
    median_mean_depth=median(mean_depth,na.rm=TRUE),
    median_1000x_percent=median(pct_1000x,na.rm=TRUE),
    .groups="drop")

timepoint_summary
```

## 12.4 Perform paired comparisons

```{r test-timepoint-qc}
timepoint_wide<-timepoint_qc %>%
  select(patient,endpoint,read_pairs_million,molecules_thousands,mean_depth,pct_1000x) %>%
  pivot_wider(names_from=endpoint,values_from=c(read_pairs_million,molecules_thousands,mean_depth,pct_1000x))

yield_test<-wilcox.test(
  timepoint_wide$read_pairs_million_BL,
  timepoint_wide$read_pairs_million_END,
  paired=TRUE,
  exact=FALSE)

molecule_test<-wilcox.test(
  timepoint_wide$molecules_thousands_BL,
  timepoint_wide$molecules_thousands_END,
  paired=TRUE,
  exact=FALSE)

depth_test<-wilcox.test(timepoint_wide$mean_depth_BL,timepoint_wide$mean_depth_END,paired=TRUE,exact=FALSE)

breadth_test<-wilcox.test(timepoint_wide$pct_1000x_BL,timepoint_wide$pct_1000x_END,paired=TRUE,exact=FALSE)

timepoint_tests<-tibble(
  measurement=c("Sequencing yield","Independent duplex molecules","Mean duplex depth","Bases covered at ≥1,000×"),
  p_value=c(
    yield_test$p.value,
    molecule_test$p.value,
    depth_test$p.value,
    breadth_test$p.value)) %>%
  mutate(adjusted_p=p.adjust(p_value,method="BH"))

timepoint_tests
```

```{r summarise-review-reasons}
review_reasons<-patient_qc %>%
  mutate(
    low_mean_depth=mean_depth<qc_limits$lower_mean_depth,
    low_500x_coverage=pct_500x<qc_limits$lower_500x,
    low_molecule_count=independent_duplex_molecules<qc_limits$lower_molecules,
    high_zero_coverage=zero_percent>qc_limits$upper_zero_percent) %>%
  filter(qc_status=="Review") %>%
  select(
    sample,
    endpoint,
    low_mean_depth,
    low_500x_coverage,
    low_molecule_count,
    high_zero_coverage,
    mean_depth,
    independent_duplex_molecules,
    zero_percent,
    pct_500x,
    pct_1000x)

review_reasons
```

## 12.5 Plot timepoint comparisons

```{r plot-timepoint-qc}
timepoint_colours<-c("Baseline"="darkcyan","Post-radiotherapy"="lightpink4")

plot_timepoint_yield<-ggplot(timepoint_qc,aes(x=timepoint,y=read_pairs_million,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_jitter(width=0.08,colour="black")+
  scale_fill_manual(values=timepoint_colours)+
  labs(x="",y="Sequencing yield\n(million read pairs)")+
  theme_unified()+
  theme(legend.position="none")

plot_timepoint_molecules<-ggplot(timepoint_qc,aes(x=timepoint,y=molecules_thousands,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_jitter(width=0.08,colour="black")+
  scale_fill_manual(values=timepoint_colours)+
  labs(x="",y="Independent duplex\nmolecules (×10³)")+
  theme_unified()+
  theme(legend.position="none")

plot_timepoint_depth<-ggplot(timepoint_qc,aes(x=timepoint,y=mean_depth,fill=timepoint))+
  geom_boxplot(width=0.5,outlier.shape=NA,alpha=0.7)+
  geom_jitter(width=0.08,colour="black")+
  scale_fill_manual(values=timepoint_colours)+
  labs(x="",y="Mean duplex-consensus\ncoverage (×)")+
  theme_unified()+
  theme(legend.position="none")

figure_timepoint_balance<-patchwork::wrap_plots(
  plot_timepoint_yield,
  plot_timepoint_molecules,
  plot_timepoint_depth,
  ncol=3)+
  patchwork::plot_annotation(tag_levels="A")

figure_timepoint_balance
```

## 12.6 Save timepoint figure

```{r save-timepoint-qc}
ggsave(file.path(figure_path,"timepoint_qc_balance.pdf"),plot=figure_timepoint_balance,width=11,height=4)

ggsave(file.path(figure_path,"timepoint_qc_balance.png"),plot=figure_timepoint_balance,width=11,height=4,dpi=600)
```

# 13 Publication figures

## 13.1 Build figures

```{r final-publication-figure-build, include=FALSE}
  #Settings
  point_size <- 0.55
  point_alpha <- 0.60
  box_width <- 0.45
  jitter_width <- 0.08
  box_linewidth <- 0.45
  line_width <- 0.75

  patient_col <- unname(unified_colours["Patient samples"])
  healthy_col <- unname(unified_colours["Healthy controls"])
  raw_col <- unname(unified_colours["Raw reads"])
  duplex_col <- unname(unified_colours["Duplex consensus"])
  ideal_col <- unname(unified_colours["Ideal duplex fraction"])

  stage_cols <- c("Raw reads"=raw_col,"Duplex consensus"=duplex_col)
  group_cols <- c("Healthy controls"=healthy_col,"Patient samples"=patient_col)
  timepoint_cols <- c("Baseline"=unname(unified_colours["Baseline"]),
                      "Post-radiotherapy"=unname(unified_colours["Post-radiotherapy"]))
  saturation_cols <- c("Observed duplex fraction"=duplex_col,"Ideal duplex fraction"=ideal_col)

  qc_theme <- theme_unified(base_size=10,base_family="sans")+
    theme(axis.title=element_text(size=10),
          axis.text=element_text(size=9,colour="black"),
          axis.title.x=element_text(margin=margin(t=5)),
          axis.title.y=element_text(margin=margin(r=6)),
          axis.line=element_line(linewidth=0.45),
          axis.ticks=element_line(linewidth=0.40),
          axis.ticks.length=grid::unit(1.2,"mm"),
          legend.title=element_text(size=9),
          legend.text=element_text(size=8.5),
          legend.key.height=grid::unit(3.5,"mm"),
          legend.key.width=grid::unit(5,"mm"),
          legend.margin=margin(0),
          legend.box.margin=margin(0),
          legend.box.spacing=grid::unit(0.5,"mm"),
          legend.spacing.x=grid::unit(2,"mm"),
          plot.margin=margin(7,7,6,7))

  legend_bottom <- theme(legend.position="bottom",
                         legend.direction="horizontal",
                         legend.justification="center",
                         legend.box.just="center",
                         legend.margin=margin(0),
                         legend.box.margin=margin(0),
                         legend.box.spacing=grid::unit(0,"mm"),
                         legend.text=element_text(size=8))

  tag_theme <- theme(plot.tag=element_text(size=11,face="bold"),
                     plot.tag.position=c(0.02,0.98),
                     plot.tag.location="panel")

  #Match single-category box width
  single_box_scale <- scale_x_discrete(expand=expansion(add=c(0.80,0.80)))
  multiple_box_scale <- scale_x_discrete(expand=expansion(add=c(0.30,0.30)))

  box_layer <- function(fill=patient_col) {
    geom_boxplot(width=box_width,outlier.shape=NA,fill=fill,colour="black",
                 alpha=0.70,linewidth=box_linewidth,show.legend=FALSE)
  }

  point_layer <- function(width=jitter_width) {
    geom_jitter(width=width,height=0,size=point_size,alpha=point_alpha,colour="black",show.legend=FALSE)
  }

  qc_figure_dir <- if(exists("figure_path") && length(figure_path)==1 &&
                      !is.na(figure_path) && nzchar(figure_path)) figure_path else "."

  dir.create(qc_figure_dir,recursive=TRUE,showWarnings=FALSE)

  save_qc <- function(plot,file,width,height) {
    ggsave(file.path(qc_figure_dir,paste0(file,".pdf")),plot=plot,width=width,height=height,useDingbats=FALSE)
    ggsave(file.path(qc_figure_dir,paste0(file,".png")),plot=plot,width=width,height=height,dpi=600,bg="white")
  }

  set.seed(123)

  #Main Figure 1
  target_gene_order <- c("BAP1","SF3B1","EIF1AX","MBD4","GNAQ","GNA11","PLCB4","CYSLTR2")

  patient_gene_plot_data <- gene_coverage %>%
    filter(sample_group=="Patient samples") %>%
    mutate(gene=factor(as.character(gene),levels=target_gene_order))

  plot_target_depth <- ggplot(patient_gene_plot_data,aes(x=gene,y=mean_gene_coverage))+
    geom_boxplot(width=0.58,outlier.shape=NA,fill=patient_col,colour="black",alpha=0.72,linewidth=box_linewidth)+
    geom_jitter(width=0.10,height=0,size=0.42,alpha=0.52,colour="black",show.legend=FALSE)+
    scale_y_continuous(labels=label_comma(),expand=expansion(mult=c(0.02,0.08)))+
    labs(x=NULL,y="Mean target coverage (x)")+
    qc_theme+
    theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1))

  patient_gene_breadth <- base_gene %>%
    filter(patient!="1452",sample_group=="Patient samples") %>%
    select(sample,patient,gene,pct_500x,pct_1000x) %>%
    pivot_longer(cols=c(pct_500x,pct_1000x),names_to="coverage_threshold",values_to="percent_covered") %>%
    mutate(
      gene=factor(gene,levels=target_gene_order),
      coverage_threshold=recode(coverage_threshold,pct_500x=">=500x",pct_1000x=">=1,000x"),
      coverage_threshold=factor(
        coverage_threshold,
        levels=c(">=500x",">=1,000x")))

  gene_breadth_summary_main <- patient_gene_breadth %>%
    group_by(gene,coverage_threshold) %>%
    summarise(
      median=median(percent_covered,na.rm=TRUE),
      Q1=quantile(percent_covered,0.25,na.rm=TRUE),
      Q3=quantile(percent_covered,0.75,na.rm=TRUE),
      .groups="drop")

  breadth_dodge <- position_dodge(width=0.52)
  breadth_cols <- c(">=500x"=patient_col,">=1,000x"="#D55E00")

  plot_target_breadth <- ggplot(gene_breadth_summary_main,aes(x=gene,y=median,colour=coverage_threshold))+
    geom_linerange(aes(ymin=Q1,ymax=Q3),position=breadth_dodge,linewidth=0.65)+
    geom_point(position=breadth_dodge,size=2.1)+
    scale_colour_manual(values=breadth_cols)+
    scale_y_continuous(
      limits=c(0,100),breaks=seq(0,100,20),
      labels=label_percent(scale=1),
      expand=expansion(mult=c(0,0.02)))+
    labs(x=NULL,y="Target bases covered (%)",colour="Coverage threshold")+
    qc_theme+legend_bottom+
    theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1))

  patient_interval_support <- interval_data %>%
    filter(endpoint%in%c("Baseline","Post-radiotherapy")) %>%
    group_by(name,interval_number) %>%
    summarise(
      patient_samples=n(),
      percent_samples_500x=100*mean(mean_coverage>=500,na.rm=TRUE),
      median_coverage=median(mean_coverage,na.rm=TRUE),
      .groups="drop") %>%
    mutate(
      name=factor(name,levels=target_gene_order),
      support_class=if_else(
        percent_samples_500x==100,
        "All samples >=500x",
        "Some samples <500x"))

  interval_support_counts <- patient_interval_support %>%
    summarise(
      intervals=n(),
      all_samples_500x=sum(percent_samples_500x==100),
      at_least_95_percent_500x=sum(percent_samples_500x>=95))

  interval_gene_support <- patient_interval_support %>%
    group_by(name) %>%
    summarise(
      intervals=n(),
      all_samples=sum(percent_samples_500x==100),
      at_least_95_percent=sum(percent_samples_500x>=95),
      .groups="drop") %>%
    pivot_longer(cols=c(all_samples,at_least_95_percent),names_to="support_threshold",values_to="passing_intervals") %>%
    mutate(
      name=factor(as.character(name),levels=target_gene_order),
      percent_intervals=100*passing_intervals/intervals,
      count_label=paste0(passing_intervals,"/",intervals),
      support_threshold=recode(support_threshold,all_samples="100% of samples",at_least_95_percent=">=95% of samples"),
      support_threshold=factor(
        support_threshold,
        levels=c("100% of samples",">=95% of samples")))

  interval_dodge <- position_dodge(width=0.72)
  interval_cols <- c("100% of samples"=patient_col,">=95% of samples"="#D55E00")

  plot_interval_support <- ggplot(interval_gene_support,aes(x=name,y=percent_intervals,fill=support_threshold))+
    geom_col(position=interval_dodge,width=0.62,colour="black",linewidth=0.30)+
    geom_text(aes(label=count_label),position=interval_dodge,vjust=-0.35,size=2.7,show.legend=FALSE)+
    scale_fill_manual(values=interval_cols)+
    scale_y_continuous(breaks=seq(0,100,20),labels=label_percent(scale=1),expand=expansion(mult=c(0,0.08)))+
    coord_cartesian(ylim=c(0,108),clip="off")+
    labs(x=NULL,y="Capture intervals meeting\n500x criterion (%)",fill="Patient-sample support")+
    qc_theme+legend_bottom+
    theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1))

  main_panel <- wrap_plots(plot_target_depth,plot_target_breadth,plot_interval_support,ncol=1,heights=c(1.05,0.95,1.0))+
    plot_annotation(
      tag_levels="A",
      caption=paste0(
        "Patient samples, n=",n_distinct(patient_gene_plot_data$sample),
        ". ",interval_support_counts$all_samples_500x,"/",
        interval_support_counts$intervals,
        " intervals reached 500x in every sample; ",
        interval_support_counts$at_least_95_percent_500x,"/",
        interval_support_counts$intervals,
        " reached 500x in at least 95% of samples."),
      theme=tag_theme+
        theme(plot.caption=element_text(size=8,hjust=0,margin=margin(t=5))))

  save_qc(main_panel,"Figure_1_Targeted_panel_coverage_QC",7.2,8.6)

  #Supplementary Figure S1
  plot_yield <- yield_data %>%
    mutate(group="Patient samples") %>%
    ggplot(aes(x=group,y=read_pairs_million))+
    box_layer()+point_layer()+single_box_scale+
    labs(x=NULL,y="Read pairs\n(millions)")+
    qc_theme+
    theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())

  plot_molecules <- molecule_data %>%
    mutate(group="Patient samples") %>%
    ggplot(aes(x=group,y=molecules_thousands))+
    box_layer()+point_layer()+single_box_scale+
    scale_y_continuous(labels=label_comma())+
    labs(x=NULL,y="Duplex molecules\n(\u00D710\u00B3)")+
    qc_theme+
    theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())

  coverage_plot_data <- coverage_data %>%
    mutate(stage=factor(stage,levels=c("Raw reads","Duplex consensus")))

  plot_coverage <- ggplot(coverage_plot_data,aes(x=stage,y=mean_target_coverage,fill=stage))+
    geom_boxplot(width=box_width,outlier.shape=NA,colour="black",alpha=0.70,linewidth=box_linewidth,show.legend=FALSE)+
    point_layer()+
    scale_fill_manual(values=stage_cols)+
    scale_x_discrete(labels=c("Raw reads"="Raw","Duplex consensus"="Duplex"),expand=expansion(add=c(0.30,0.30)))+
    scale_y_log10(breaks=c(1000,10000,100000),labels=c("1k","10k","100k"))+
    labs(x=NULL,y="Target coverage\n(\u00D7; log scale)")+
    qc_theme

  plot_saturation <- ggplot(saturation_summary,aes(x=sequencing_fraction))+
    geom_ribbon(aes(ymin=observed_Q1,ymax=observed_Q3),fill=duplex_col,alpha=0.15)+
    geom_ribbon(aes(ymin=ideal_Q1,ymax=ideal_Q3),fill=ideal_col,alpha=0.10)+
    geom_line(aes(y=observed_median,colour="Observed duplex fraction"),linewidth=line_width)+
    geom_line(aes(y=ideal_median,colour="Ideal duplex fraction"),linewidth=line_width,linetype="dashed")+
    scale_colour_manual(values=saturation_cols)+
    scale_x_continuous(limits=c(0,100),breaks=seq(0,100,20),expand=expansion(mult=c(0,0.02)))+
    scale_y_continuous(limits=c(0,100),breaks=seq(0,100,20),
                       labels=label_percent(scale=1),
                       expand=expansion(mult=c(0,0.02)))+
    labs(x="Sequencing data used (%)",y="Duplex fraction (%)",colour=NULL)+
    qc_theme+legend_bottom

  plot_breadth <- ggplot(breadth_summary,aes(x=coverage_threshold,group=stage))+
    geom_ribbon(aes(ymin=lower_IQR,ymax=upper_IQR,fill=stage),alpha=0.15,colour=NA,show.legend=FALSE)+
    geom_line(aes(y=median,colour=stage),linewidth=line_width)+
    geom_point(aes(y=median,colour=stage),size=point_size)+
    scale_colour_manual(values=stage_cols)+
    scale_fill_manual(values=stage_cols)+
    scale_x_log10(breaks=c(10,20,50,100,250,500,1000),labels=c("10","20","50","100","250","500","1k"))+
    scale_y_continuous(limits=c(0,100),breaks=seq(0,100,20),
                       labels=label_percent(scale=1),
                       expand=expansion(mult=c(0,0.02)))+
    labs(x="Coverage threshold (\u00D7)",y="Target bases\ncovered (%)",colour=NULL)+
    qc_theme+legend_bottom

  #Align panels in separate rows
  supp_s1_top <- wrap_plots(plot_yield,plot_molecules,plot_coverage,ncol=3,widths=c(1,1,1))

  supp_s1_bottom <- wrap_plots(plot_saturation,plot_breadth,ncol=2,widths=c(1,1))

  supp_s1 <- wrap_plots(supp_s1_top,supp_s1_bottom,ncol=1,heights=c(1,1.15))+
    plot_annotation(tag_levels="A",theme=tag_theme)

  save_qc(supp_s1,"Supplementary_Figure_S1_Sequencing_and_consensus_performance",7.2,6.2)

  #Supplementary Figure S2
  plot_uniformity <- ggplot(uniformity_paired,aes(x=raw,y=duplex))+
    geom_abline(slope=1,intercept=0,linetype="dashed",colour="grey50",linewidth=0.60)+
    geom_point(size=point_size,alpha=point_alpha,colour=duplex_col)+
    coord_equal(xlim=c(1,1.65),ylim=c(1,1.65),expand=FALSE)+
    scale_x_continuous(breaks=seq(1,1.6,0.2))+
    scale_y_continuous(breaks=seq(1,1.6,0.2))+
    labs(x="Raw-read Fold-80\nbase penalty",y="Duplex-consensus\nFold-80 base penalty")+
    qc_theme

  plot_duplex_fraction <- ggplot(duplex_data,aes(x=duplex_fraction,y=ideal_duplex_fraction))+
    geom_abline(slope=1,intercept=0,linetype="dashed",colour="grey50",linewidth=0.60)+
    geom_point(size=point_size,alpha=point_alpha,colour=duplex_col)+
    coord_equal(xlim=c(50,102),ylim=c(50,102),expand=FALSE)+
    scale_x_continuous(breaks=seq(50,100,10),labels=label_percent(scale=1))+
    scale_y_continuous(breaks=seq(50,100,10),labels=label_percent(scale=1))+
    labs(x="Observed duplex\nfraction (%)",y="Ideal duplex\nfraction (%)")+
    qc_theme

  supp_s2 <- wrap_plots(plot_uniformity,plot_duplex_fraction,ncol=2,widths=c(1,1))+
    plot_annotation(tag_levels="A",theme=tag_theme)

  save_qc(supp_s2,"Supplementary_Figure_S2_Fold80_and_duplex_fraction",7.2,3.9)

  #Supplementary Figure S3
  plot_family_size <- ggplot(duplex_family_size_summary,aes(x=family_size_category,y=median_percent))+
    geom_col(width=0.68,fill=patient_col,colour="black",linewidth=0.35,alpha=0.80)+
    geom_errorbar(aes(ymin=Q1,ymax=Q3),width=0.16,linewidth=0.50)+
    geom_text(aes(y=Q3,label=paste0(sprintf("%.1f",median_percent),"%")),vjust=-0.45,size=3.2)+
    scale_y_continuous(labels=label_percent(scale=1),expand=expansion(mult=c(0,0.14)))+
    labs(
      x="Minimum reads in each strand family",
      y="Duplex families per sample (%)",
      subtitle="Bars show medians; whiskers show the interquartile range",
      caption=paste0(
        "Categories use BA, the smaller strand family; ",
        "the larger AB family may contain additional reads."))+
    qc_theme+
    theme(plot.subtitle=element_text(size=9),plot.caption=element_text(size=8,hjust=0,margin=margin(t=5)))

  save_qc(plot_family_size,"Supplementary_Figure_S3_UMI_family_size_distribution",6.2,4.8)

  #Supplementary Figure S4
  gene_plot_data <- gene_coverage %>%
    mutate(gene=factor(gene,levels=c("BAP1","SF3B1","EIF1AX","MBD4",
                                     "GNAQ","GNA11","PLCB4","CYSLTR2")),
           sample_group=factor(sample_group,
                               levels=c("Healthy controls","Patient samples")))

  plot_gene <- ggplot(gene_plot_data,aes(x=gene,y=mean_gene_coverage,fill=sample_group))+
    geom_boxplot(position=position_dodge(width=0.70),width=0.55,
                 outlier.shape=NA,colour="black",alpha=0.70,
                 linewidth=box_linewidth)+
    geom_point(aes(group=sample_group),
               position=position_jitterdodge(jitter.width=0.06,dodge.width=0.70),
               size=point_size,alpha=point_alpha,colour="black",
               show.legend=FALSE)+
    scale_fill_manual(values=group_cols)+
    scale_y_continuous(labels=label_comma())+
    labs(x="Targeted gene",y="Duplex-consensus\ncoverage (\u00D7)",fill=NULL)+
    qc_theme+legend_bottom+
    theme(axis.text.x=element_text(angle=30,hjust=1,vjust=1))

  plot_base_depth <- ggplot(base_threshold_summary,aes(x=threshold,group=sample_group))+
    geom_ribbon(aes(ymin=Q1,ymax=Q3,fill=sample_group),alpha=0.16,colour=NA,show.legend=FALSE)+
    geom_line(aes(y=median,colour=sample_group),linewidth=line_width)+
    geom_point(aes(y=median,colour=sample_group),size=point_size)+
    scale_colour_manual(values=group_cols)+
    scale_fill_manual(values=group_cols)+
    scale_x_discrete(labels=c("250\u00D7","500\u00D7","1,000\u00D7","2,000\u00D7","3,000\u00D7"))+
    scale_y_continuous(limits=c(0,100),breaks=seq(0,100,20),
                       labels=label_percent(scale=1),
                       expand=expansion(mult=c(0,0.02)))+
    labs(x="Coverage threshold (\u00D7)",y="Target bases\ncovered (%)",colour=NULL)+
    qc_theme+legend_bottom

  supp_s4 <- wrap_plots(plot_gene,plot_base_depth,ncol=1,heights=c(1.05,0.95))+
    plot_annotation(tag_levels="A",theme=tag_theme)

  save_qc(supp_s4,"Supplementary_Figure_S4_Gene_and_base_coverage_by_group",7.2,7.8)

  #Supplementary Figure S5
  publication_interval_summary <- interval_data %>%
    filter(endpoint%in%c("Baseline","Post-radiotherapy")) %>%
    group_by(name,interval_number) %>%
    summarise(
      patient_samples=n(),
      minimum=min(mean_coverage,na.rm=TRUE),
      Q1=quantile(mean_coverage,0.25,na.rm=TRUE),
      median=median(mean_coverage,na.rm=TRUE),
      Q3=quantile(mean_coverage,0.75,na.rm=TRUE),
      maximum=max(mean_coverage,na.rm=TRUE),
      percent_samples_500x=100*mean(mean_coverage>=500,na.rm=TRUE),
      .groups="drop") %>%
    mutate(
      name=factor(name,levels=target_gene_order),
      interval_label=factor(interval_number),
      support_class=case_when(
        percent_samples_500x==100 ~ "100% of samples",
        percent_samples_500x>=95 ~ ">=95% of samples",
        TRUE ~ "<95% of samples"),
      support_class=factor(
        support_class,
        levels=c("100% of samples",">=95% of samples","<95% of samples")))

  interval_support_cols <- c("100% of samples"=patient_col,">=95% of samples"="#D55E00","<95% of samples"="#CC79A7")

  make_interval_summary_plot <- function(data,y_title=TRUE) {
    ggplot(data,aes(x=median,y=interval_label,colour=support_class))+
      geom_vline(xintercept=500,linetype="dashed",colour="grey45",linewidth=0.55)+
      geom_segment(aes(x=minimum,xend=maximum,y=interval_label,yend=interval_label),colour="grey72",linewidth=0.35)+
      geom_segment(aes(x=Q1,xend=Q3,y=interval_label,yend=interval_label),linewidth=1.05)+
      geom_point(size=1.7)+
      facet_grid(rows=vars(name),scales="free_y",space="free_y",switch="y")+
      scale_colour_manual(
        values=interval_support_cols,
        breaks=c("100% of samples",">=95% of samples","<95% of samples"),
        labels=c("100%","\u226595%","<95%"),
        drop=FALSE)+
      scale_x_log10(
        breaks=c(250,500,1000,2000,5000),
        labels=c("250","500","1k","2k","5k"),
        expand=expansion(mult=c(0.03,0.08)))+
      scale_y_discrete(limits=function(x) rev(x))+
      labs(
        x="Interval coverage (\u00D7; log scale)",
        y=if(y_title) "Capture interval" else NULL,
        colour="Samples \u2265500\u00D7")+
      qc_theme+
      theme(
        text=element_text(size=10),
        axis.title=element_text(size=10),
        axis.text=element_text(size=10,colour="black"),
        legend.position="none",
        strip.placement="outside",
        strip.background=element_blank(),
        strip.text.y.left=element_text(angle=0,size=10,face="bold"),
        panel.spacing.y=grid::unit(0.7,"mm"),
        plot.margin=margin(5,5,4,13))
  }

  plot_intervals_sf3b1 <- make_interval_summary_plot(publication_interval_summary %>% filter(name=="SF3B1"))

  plot_intervals_other <- make_interval_summary_plot(
    publication_interval_summary %>% filter(name!="SF3B1"),
    y_title=FALSE)

  plot_intervals <- (plot_intervals_sf3b1+plot_intervals_other+ plot_layout(ncol=2,widths=c(1,1)))+
    plot_annotation(
      caption=paste0(
        "Points show medians; thick lines show IQRs; thin lines show ranges.\n",
        "Colours show samples reaching \u2265500\u00D7: ",
        "teal, 100%; orange, \u226595%; pink, <95%."),
      theme=theme(
        plot.caption=element_text(size=10,hjust=0,margin=margin(t=3))))&
    theme(text=element_text(size=10))

  save_qc(plot_intervals,"Supplementary_Figure_S5_Target_interval_coverage_summary",7.2,6.4)

  #Supplementary Figure S6
  timepoint_data <- timepoint_qc %>%
    mutate(timepoint=factor(timepoint,levels=c("Baseline","Post-radiotherapy")))

  make_timepoint_plot <- function(data,y,ylabel,comma=FALSE) {
    p <- ggplot(data,aes(x=timepoint,y={{y}},fill=timepoint))+
      geom_boxplot(width=box_width,outlier.shape=NA,colour="black",
                   alpha=0.70,linewidth=box_linewidth,show.legend=FALSE)+
      point_layer()+
      scale_fill_manual(values=timepoint_cols)+
      scale_x_discrete(labels=c("Baseline"="Baseline",
                                "Post-radiotherapy"="Post-\nradiotherapy"),
                       expand=expansion(add=c(0.30,0.30)))+
      labs(x=NULL,y=ylabel)+
      qc_theme

    if(comma) p <- p+scale_y_continuous(labels=label_comma())
    p
  }

  plot_tp_yield <- make_timepoint_plot(timepoint_data,read_pairs_million,"Read pairs\n(millions)")

  plot_tp_molecules <- make_timepoint_plot(timepoint_data,molecules_thousands,"Duplex molecules\n(\u00D710\u00B3)",TRUE)

  plot_tp_depth <- make_timepoint_plot(timepoint_data,mean_depth,"Duplex-consensus\ncoverage (\u00D7)",TRUE)

  supp_s6 <- wrap_plots(plot_tp_yield,plot_tp_molecules,plot_tp_depth,ncol=3,widths=c(1,1,1))+
    plot_annotation(tag_levels="A",theme=tag_theme)

  save_qc(supp_s6,"Supplementary_Figure_S6_Endpoint_sequencing_QC",7.2,3.9)
```

## 13.2 Main targeted-panel QC figure

```{r main-targeted-panel-qc, echo=FALSE, fig.width=7.2, fig.height=8.6, dpi=150}
main_panel
```

## 13.3 Supplementary figures

```{r supplementary-figures, echo=FALSE, out.width='100%', fig.align='center'}
supplementary_figure_files <- file.path(
  qc_figure_dir,
  c(
    "Supplementary_Figure_S1_Sequencing_and_consensus_performance.png",
    "Supplementary_Figure_S2_Fold80_and_duplex_fraction.png",
    "Supplementary_Figure_S3_UMI_family_size_distribution.png",
    "Supplementary_Figure_S4_Gene_and_base_coverage_by_group.png",
    "Supplementary_Figure_S5_Target_interval_coverage_summary.png",
    "Supplementary_Figure_S6_Endpoint_sequencing_QC.png"))

stopifnot(all(file.exists(supplementary_figure_files)))
knitr::include_graphics(supplementary_figure_files)
```
